### 2025-07-01 docs/paper.tex
write theoretical background (recent refinements - credit constraints)

% AER-Article.tex for AEA last revised 22 June 2011
\documentclass[]{AEA}

% The mathtime package uses a Times font instead of Computer Modern.
% Uncomment the line below if you wish to use the mathtime package:
%\usepackage[cmbold]{mathtime}
% Note that miktex, by default, configures the mathtime package to use commercial fonts
% which you may not have. If you would like to use mathtime but you are seeing error
% messages about missing fonts (mtex.pfb, mtsy.pfb, or rmtmi.pfb) then please see
% the technical support document at http://www.aeaweb.org/templates/technical_support.pdf
% for instructions on fixing this problem.

% Note: you may use either harvard or natbib (but not both) to provide a wider
% variety of citation commands than latex supports natively. See below.

% Uncomment the next line to use the natbib package with bibtex
\usepackage{natbib}

% Uncomment the next line to use the harvard package with bibtex
%\usepackage[abbr]{harvard}

% This command determines the leading (vertical space between lines) in draft mode
% with 1.5 corresponding to "double" spacing.
\draftSpacing{1.5}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Add imagehandling

\usepackage{hyperref}

\begin{document}

\title{Nutritional Deficits and the Quantity-Quality Trade-off: Evidence
from an Exogenous Fertility Shock in Low-Income Urban Settings in the
Philippines}
\shortTitle{The Q-Q Trade-off in the Philippine Context}
% \author{Author1 and Author2\thanks{Surname1: affiliation1, address1, email1.
% Surname2: affiliation2, address2, email2. Acknowledgements}}

\author{
  Erika Salvador\\
  Caroline Theoharides\thanks{
  Salvador: Amherst
College, \href{mailto:esalvador28@amherst.edu}{esalvador28@amherst.edu}.
  Theoharides: Amherst
College, \href{mailto:ctheoharides@amherst.edu}{ctheoharides@amherst.edu}.
  This research was made possible by the Schupf Fellowship at Amherst
  College. I am indebted to my faculty adviser, Professor Caroline
  Theoharides, for her mentorship and support throughout the development
  of this project. I also wish to thank Faculty Director Professor
  Amelie Hastie and the campus partners whose institutional support
  enabled this work. I am further grateful to the Departments of
  Economics and Mathematics \& Statistics for their academic support.
  The views expressed and any errors contained herein are entirely my
  own responsibility.
}
}

\date{\today}
\pubMonth{06}
\pubYear{2025}
\pubVolume{}
\pubIssue{}
\JEL{J13, I15, O15}
\Keywords{fertility shocks, child nutrition, quantity-quality
trade-off, urban poverty, Philippines}

\begin{abstract}
This paper examines whether increased fertility affects early-life
nutritional outcomes in low-income urban households. I exploit a natural
experiment created by a 1990 policy in Manila, Philippines, which banned
modern contraceptives from city-run health facilities. Using a
difference-in-differences framework and nationally representative data
from the Philippine Demographic and Health Surveys, I estimate the
reduced-form impact of the policy on child height-for-age and
weight-for-height. {[}Will add more after data analysis{]}
\end{abstract}

\maketitle

\section{Introduction}

The trade-off between child quantity and child quality is a foundational
concept in the economics of the family. First articulated by
\citet{becker1960economic} and extended in subsequent models of
household behavior \citep{becker1973interaction, becker1976child}, this
framework posits that parents allocate finite resources---both financial
and non-financial---across children. An increase in fertility reduces
the resources available per child and, under binding constraints, may
lead to lower investments in health, education, and other forms of human
capital. This mechanism has served as an explanatory model for changes
in fertility behavior and the evolution of population structures in low-
and middle-income countries.

Empirical investigations of the quantity--quality trade-off have focused
primarily on educational outcomes. Studies in both high-income and
low-income settings have examined the effects of fertility on school
enrollment, grade progression, test scores, and completed years of
schooling
\citep{rosenzweig1980testing, black2005more, angrist2010effects}. These
outcomes serve as accessible proxies for long-run human capital
accumulation, but they represent only one dimension of child quality.
Other outcomes, such as nutritional status, are early-onset or
biologically constrained. They also tend to be less responsive to
remediation later in life. A child who suffers from chronic malnutrition
may exhibit permanently reduced cognitive capacity and face limits in
physical development that affect long-run productivity regardless of
subsequent educational access
\citep{hoddinott2013adult, grantham2007development}.

The exclusion of nutritional outcomes from much of the empirical
literature leaves an important dimension of the trade-off untested.
Nutritional investments in early childhood are essential to early
childhood development and long-term outcomes
\citep{victora2008maternal, hoddinott2013adult}. They shape brain
development, immune system functioning, and physical stature, and they
have been shown to predict later-life earnings and health outcomes
across a wide range of settings
\citep{grantham2007development, alderman2006long}. The biological
irreversibility of early-life nutritional deficits further distinguishes
them from other forms of investment. Educational deficits may be
partially remediable; nutritional failures often are not. A credible
estimate of the trade-off between fertility and child quality must
account for nutrition if it aims to assess the full set of consequences
associated with fertility shocks.

This study addresses this gap by examining the nutritional effects of a
localized, exogenous increase in fertility in the Philippines. In 1990,
the mayor of Manila implemented an executive policy that prohibited the
provision of modern contraceptives in all city-run health facilities.
The order removed access to pills, condoms, intrauterine devices, and
related public health materials and instructed healthcare providers to
offer only natural family planning methods. This policy remained in
place for nearly a decade and affected only the jurisdiction of the
Manila city government. The national government did not implement a
comparable restriction, and surrounding cities within Metro Manila
continued to provide access to modern contraceptives. The policy thus
created a spatial and temporal discontinuity in contraceptive access
that was uncorrelated with underlying fertility preferences or
concurrent shifts in household income or governance. As a result, the
Manila ban serves as a quasi-experimental source of variation in
fertility exposure among poor urban households.

I use this natural experiment to estimate the causal effect of increased
fertility on child nutrition. The analysis relies on nationally
representative data from multiple waves of the Philippine Demographic
and Health Survey (DHS), which provide data on household structure and
maternal characteristics, as well as measurements of child
anthropometry. The outcomes of interest are height-for-age and
weight-for-height z-scores, which serve as standardized indicators of
chronic and acute malnutrition, respectively. These outcomes are widely
used in the global health and development literature and capture
nutritional deprivation over both long and short time horizons
\citep{victora2008maternal}. The empirical strategy follows a
difference-in-differences design that compares child outcomes in Manila
and comparable urban areas before and after the onset of the policy.

The identification strategy rests on two key assumptions. First, in the
absence of the contraceptive ban, nutritional trends in Manila would
have evolved in parallel with those in comparison cities. Second, any
other policy or economic shocks affecting Manila during the study period
must not coincide precisely with the timing and scope of the
contraceptive policy. I test these assumptions using falsification
checks, placebo comparisons, and robustness specifications that include
city-specific time trends, maternal fixed effects, and controls for
baseline demographic differences.

The analysis proceeds in three stages. I first replicate existing work
\citep{dumas2019sex} to confirm that the contraceptive ban led to an
increase in fertility among affected women. I then estimate reduced-form
effects of policy exposure on nutritional outcomes for children under
five years of age. Finally, I examine heterogeneity in effects across
subsamples defined by maternal education, household wealth, and access
to prenatal care. These dimensions serve as proxies for household
resource availability and capacity to buffer the nutritional
consequences of fertility increases.

This study contributes to the literature in several important ways. It
provides new evidence on how increases in fertility---caused by policy
restrictions on family planning---can affect child nutrition in poor,
urban communities. Most past research has focused on education, but this
study expands the idea of child quality to include biological measures
such as stunting and wasting. It also adds to the small number of
studies that use unexpected changes in reproductive health policy to
examine long-term effects on children's well-being. More broadly, the
results show that local restrictions on family planning can
unintentionally harm children's health, especially in settings where
families already face poverty, food insecurity, and limited public
services.

\section{Review of Related Literature}

The quantity--quality (Q--Q) theory, a central idea in modern family
economics, holds that parents face a trade-off between the number of
children and the ``quality'' of investment---such as education or
health---they can provide to each. Quality in this context refers to the
human capital of each child: attributes like education, health, and
nutrition that enhance a child's future productivity and well-being. The
genesis of this idea traces back to Gary Becker's seminal work around
1960, which for the first time treated children as economic goods
subject to parental choice and budget constraints
\citep{becker1960economic}. Becker argued that as families become
wealthier, they may not simply want more children, but rather
better-raised children, much as a household might prefer a
higher-quality car or house over a greater quantity of them. This
proposition led to a formal theory in which increases in income or
changes in economic conditions cause parents to substitute child quality
for quantity, consistent with historical patterns of lower fertility and
higher educational attainment during economic development
\citep{galor2000population}.

In what follows, I review the theoretical foundations of the Q-Q model
and its evolution in the literature. I begin with the static models of
Becker \citep{becker1960economic} and Becker--Lewis
\citep{becker1973interaction}, which first formalized the trade-off
within a household utility maximization framework. We then examine
extensions to dynamic, intergenerational settings, including the
contributions of Becker and Tomes \citep{becker1976child} on child
endowments and the altruistic dynastic model associated with Barro and
Becker \citep{barro1989fertility}. Next, I turn to macroeconomic and
unified growth models, notably Galor and Weil
\citep{galor2000population} and Galor and Moav \citep{galor2002natural},
which integrate the Q--Q mechanism into a general theory of demographic
and economic transformation.

Finally, I discuss more recent refinements that enrich the basic model
by incorporating credit constraints \citep{doepke2004accounting},
intra-household bargaining \citep{doepke2019bargaining}, and
multi-dimensional child quality
\citep{hoddinott2013economic, kalemli2002does}, with a special emphasis
on health and nutrition. The literature review focuses on how the Q--Q
framework has been applied to understand fertility and child investment
patterns, especially in developing country contexts where resource
constraints and health outcomes are paramount.

\subsection{Theoretical Background}

\subsubsection{Becker’s Static Model}

Becker's early work introduced an economic model of fertility and
treated children as durable goods that provide utility to parents but
impose costs \citep{doepke2015gary}. In Becker's 1960 model, a household
derives satisfaction from the number of children (\(n\)) and from the
quality of each child (\(q\)), alongside conventional consumption of
other goods (\(y\)). A simple representation is a utility function:

\[
U = U(n, q, y),
\]

with \(U\) increasing in each argument up to some satiation point. Here,
quality \(q\) can be thought of as the expenditure or investment per
child (e.g.~education spending, health care, nutrition), assumed for now
to be the same for each child. Parents face a budget constraint that
links quantity and quality: raising more children dilutes the resources
available per child. A prototypical budget constraint (in static form)
can be written as:

\[
p_y y + p_n n + p_q n q = I,
\]

where \(I\) is total family income (or full income), \(p_n\) represents
baseline, non-discretionary costs associated with each additional child
(e.g.~expenditures on food, shelter, or clothing that are incurred
irrespective of quality-enhancing investments), and \(p_q\) denotes the
marginal cost of investing in one unit of quality per child.The term
\(p_q n q\) captures total expenditure on quality for all children and
is linear in \(n\). As the number of children rises, parents must extend
any chosen level of \(q\) across a broader base, which amplifies the
total cost of quality. Conversely, the term \(p_n n\) implies that the
cost of an additional child rises with the quality level \(q\) already
chosen, since each child must meet a higher standard of care or
investment. For instance, a household that chooses to provide more
education or better health care per child incurs an additional burden
when it expands family size, as each child must receive the same
enhanced level of investment. Similarly, a larger family increases the
cumulative cost of quality, even if \(q\) remains fixed, due to the need
to replicate expenditures across more children. In short, the shadow
price of child quality increases with \(n\), and the shadow price of
child quantity increases with \(q\). The cost structure induces a mutual
dependence between quantity and quality, such that any adjustment along
one dimension alters the effective cost of the other.

Becker and Lewis (1973) formalize the mutually reinforcing nature of the
quantity--quality cost structure. An increase in \(n\) raises the total
cost required to sustain a given level of \(q\) for each child, while a
higher level of \(q\) raises the marginal cost associated with having an
additional child. For example, allocating more resources to education or
health per child increases the financial burden of expanding family
size. This interdependence links the two decisions directly. The
household cannot choose \(n\) and \(q\) in isolation; each choice alters
the marginal cost of the other.

Mathematically, the trade-off appears in the first-order conditions of
the household's optimization problem. Let \(\lambda\) represent the
Lagrange multiplier on the full-income constraint.

\[
\mathcal{L} = U(n, q, y) + \lambda \left( I - p_y y - p_n n - p_q n q \right).
\]

The first-order conditions are:

\[
\frac{\partial \mathcal{L}}{\partial n} = U_n - \lambda(p_n + p_q q) = 0, \quad
\frac{\partial \mathcal{L}}{\partial q} = U_q - \lambda p_q n = 0, \quad
\frac{\partial \mathcal{L}}{\partial y} = U_y - \lambda p_y = 0.
\]

Combining the first two yields:

\[
\frac{U_n}{U_q} = \frac{p_n + p_q q}{p_q n}.
\] This condition equates the marginal rate of substitution between
quantity and quality to the ratio of their full marginal costs. The
numerator rises with \(q\), and the denominator rises with \(n\). As one
choice increases, the relative cost of the other becomes higher. This
relationship induces substitution toward the less costly dimension. The
trade-off between quantity and quality arises from the structure of the
budget itself. It does not rely on specific assumptions about utility
curvature or intrinsic substitutability \citep{becker1973interaction}.

This formulation implies two core predictions. Firstly, although both
child quantity and child quality may rise with income, the household's
budget constraint can generate a negative relationship between income
and fertility. As income increases, total spending on children tends to
rise, but the allocation often favors quality over quantity. Becker
illustrated this with the analogy of durable goods: wealthier households
tend to upgrade the quality of a house or a car rather than acquire
additional units. In a similar way, higher-income families often direct
additional resources toward education, nutrition, or health per child.
Within the model, an income increase (\(\mathrm{d}I > 0\)) produces a
direct effect that makes children more affordable and an indirect effect
that discourages fertility. As \(q\) rises, the shadow price of an
additional child also rises. If the marginal utility from higher quality
exceeds that from larger family size, then the substitution effect
outweighs the income effect, leading to a lower optimal \(n\). This
mechanism offers a structural explanation for the demographic
transition: fertility tends to fall as households become richer, even
when preferences remain unchanged.

Furthermore, a similar logic applies to changes in the cost parameters
\(p_q\) and \(p_n\). A decline in \(p_q\), such as through a policy that
lowers the price of education or health care, increases \(q\) and raises
the marginal cost of quantity. This effect reduces optimal fertility. A
rise in \(p_n\), which may reflect higher child-rearing costs or a
greater opportunity cost of parental time, reduces the appeal of larger
families and can shift resources toward child quality. These outcomes
follow from the structure of the budget constraint, without requiring
any explicit preference for quality over quantity. Becker and Lewis
noted that these comparative static results align with observed
patterns. For example, increases in women's wages often reduce fertility
more than they reduce educational spending per child. This asymmetry
reflects the model's central feature: quantity and quality are linked
through their cost structure. An increase in one raises the marginal
cost of the other. The model explains how households make trade-offs
between the number of children and investments in each.

\subsubsection{Intergenerational Models: Altruism and Child Endowments}

While the early Q--Q models were static (one-period) representations,
subsequent contributions extended the framework to consider fertility
and child investment over multiple periods or even multiple generations.
The main development in this literature was the incorporation of
intergenerational human capital dynamics, where parents derive utility
not only from the number and quality of children in the present, but
also from the long-run outcomes of their offspring. These extensions
allowed child quality to evolve endogenously across time, rather than
being determined solely within a single period.

One of the earliest and most influential models of this kind was
proposed by \citet{becker1976child}. In their formulation, each child
enters the world with an exogenous endowment \(E\), which may reflect
factors such as cognitive ability (e.g., measured IQ or language
acquisition speed), early health status (e.g., birth weight or incidence
of neonatal complications), genetic predispositions (e.g., risk for
chronic illness, temperament, or neurodevelopmental traits), or family
background characteristics (e.g., parental education, household
stability, or neighborhood conditions). Furthermore, parents can augment
this endowment by investing resources \(q\) in the form of education,
nutrition, and other quality-enhancing inputs. The effective adult human
capital of the child might be expressed as \(H = E + f(q)\) (in a simple
additive form) or a multiplicative variant \(H = E \cdot f(q)\), where
\(f(q)\) is an increasing concave function of parental investment.

\citet{becker1976child} emphasized that variation in endowments \(E\)
can shape how parents allocate investments \(q\) across children. When
the productivity of investment increases with endowment, parents may
concentrate resources on children with higher \(E\), who are more likely
to convert additional investment into future success. In other cases,
parents may attempt to compensate for lower endowments by directing
greater investment toward disadvantaged children. Put simply,
child-specific variation in initial conditions affects not only outcomes
but also strategic parental choices. Because of this, the relationship
between income and the demand for child quality is not uniform. The
income elasticity of demand for quality may differ across households,
depending on the distribution of endowments within the family and across
the broader population.

Furthermore, \citet{becker1976child} showed that at low income levels,
much of what constitutes child quality comes from exogenous endowments,
i.e., factors like public education, neighborhood environment, or access
to basic healthcare that are not privately purchased. In these settings,
small increases in parental income may not lead to significant changes
in fertility or investment behavior. Since most of the child's future
outcomes are determined by the fixed endowment component, marginal
investment plays a smaller role. However, as income rises, private
resources become a larger part of what determines quality, and the
classic Q--Q trade-off begins to shape behavior. Parents begin to
allocate more income toward fewer children in order to enhance quality
through direct investment. Under certain theoretical conditions, such as
equal utility elasticities for quantity and quality, this framework
produces a non-monotonic relationship between income and fertility.
Fertility may decline as income rises at first, which reflects the
desire to invest more intensively per child. However, beyond some point,
once the marginal return to investment begins to flatten or saturate
relative to the fixed endowment, fertility may increase again.

The U-shaped prediction emerges only under specific assumptions, and its
validity depends on both the shape of the utility function and how
endowments relate to parental background. More broadly,
\citet{becker1976child} enriched the Q--Q framework by incorporating
elements that reflect real-world variation. They argued that not all
differences in child outcomes are the result of deliberate parental
choice. Random factors, biological traits, and socioeconomic settings
play a role. In this light, public policies, such as subsidized
schooling, early childhood programs, or universal healthcare, can
influence private fertility and investment decisions by shifting the
effective value of \(E\) across the population. If government programs
raise the floor for child endowments, then even low-income parents can
achieve better outcomes without large private sacrifices. These
policy-induced shifts in \(E\) alter the perceived return to having more
children or investing more per child.

\citet{becker1976child} also considered the possibility that endowment
is not randomly assigned but may vary systematically with income.
Higher-income households may produce children with higher \(E\) due to
better maternal nutrition, access to prenatal care, lower exposure to
environmental risk, or assortative matching on traits associated with
educational or occupational success. In these families, not only are the
resources available for investment greater, but the potential gains from
investment may also be higher, because children are better positioned to
benefit from those inputs. This interaction deepens the divide between
high- and low-income households, making it harder for disadvantaged
families to catch up. As a result, inequality can persist or even widen
across generations.

Finally, the model provides a mechanism for understanding how
imperfections in credit markets can lead to persistent disadvantages. If
parents with low income and low-\(E\) children cannot borrow to finance
quality-enhancing investment, then the next generation begins life with
the same disadvantage. Without external intervention or structural
change, this loop continues, which results in a pattern where poor
families remain poor and rich families accumulate further advantage. The
Becker-Tomes framework thus connects household-level decisions to bigger
questions about the intergenerational transmission of human capital.

Parallel to Becker and Tomes's static analysis of endowments, another
strand of the literature developed a fully dynamic version of the Q-Q
model by incorporating parental altruism toward children's welfare. In
this framework, introduced by \citet{barro1989fertility}, parents care
not only about the number and quality of their children but also about
the utility their descendants will enjoy in the future. Altruism in this
context means that parents treat their children's utility as part of
their own, thus extending the household's objective across generations.
For example, a parent may reduce personal consumption to pay for a
child's schooling, motivated not just by the child's immediate benefit
but by the satisfaction the parent gains from the child's long-term
success. This leads to a formulation of dynastic utility, where the
household's objective spans infinitely many periods and takes the form
of a recursive altruistic structure. A representative formulation is

\[
U_{0} = \sum_{t=0}^{\infty} \beta^{t}\,u(c_{t}, n_{t}),
\]

where \(c_t\) denotes the consumption of the \(t\)-th generation,
\(n_t\) the number of children, and \(\beta \in (0,1)\) the
intertemporal discount factor. Given this structure, having an extra
child \(n_t\) enters utility positively, but each child is assumed to
receive the same utility as the parent if raised at a comparable
standard of living. As a result, parents confront an intertemporal
trade-off: having more children expands the number of future utility
streams but also stretches current resources, since each child requires
support. This trade-off gives rise to an Euler equation for optimal
fertility choice, analogous to an optimal growth condition.

An implication of the dynastic model is that fertility decisions are
sensitive to macroeconomic conditions, such as the interest rate or the
rate of return on capital. A rise in interest rates increases the
opportunity cost of channeling resources into children rather than
saving, which tends to reduce current fertility---a substitution effect
across generations. At the same time, higher returns make future
generations wealthier, and this anticipated prosperity enters the
utility calculations of parents in more complex ways.
\citet{barro1989fertility} demonstrated that the model can account for
observed fertility responses to economic fluctuations and policy
interventions It can also explain historical phenomena such as postwar
baby booms and subsequent fertility declines through shifts in returns
or labor‐market opportunities.

In many dynastic models, child quality appears indirectly, often through
the child's future human capital or income. One variant assumes parents
value the aggregate human‐capital stock of their children. This
specification, combined with altruism, produces a similar trade‐off:
concentrating resources in fewer children raises each child's human
capital, which raises the dynasty's long‐run welfare. These
intergenerational extensions link micro‐level fertility decisions to
macroeconomic outcomes. By the late 1980s, work by Becker, Barro, and
others had recast fertility as an endogenous choice that interacts with
capital accumulation, income distribution, and policy. This laid the
foundation for unified growth theories, which view the quantity--quality
mechanism as central to demographic transition and long-run development.

\subsubsection{Unified Growth Models}

The unified growth theory, developed in the late 1990s and 2000s
(notably by Oded Galor and co-authors), seeks to explain in one
framework the entire sweep of economic development -- from Malthusian
stagnation, through the demographic transition, to modern growth. A
central puzzle it addresses is why fertility rates, which were
historically high and invariant to income in the Malthusian era, began
to decline sharply in tandem with industrialization and rising incomes,
eventually stabilizing at much lower levels in developed economies. The
Q--Q trade-off provides a key part of the answer in these models.
\citet{galor2000population} and \citet{galor2002natural} explicitly
incorporate parental choices about the quantity and quality of children
and show how changes in the economic environment alter those choices and
trigger demographic transitions.

In \citet{galor2000population}'s model, for instance, technological
progress gradually increases the return to human capital, especially in
skilled occupations. In the early stages, when production relies on
basic tools and techniques, unskilled labor holds more value. Under
these conditions, parents have little reason to invest in formal
schooling. Children are expected to contribute economically through
agricultural work, domestic tasks, or low-skill jobs in workshops and
factories. Fertility remains high because children impose a low
financial burden and generate immediate returns. As technology becomes
more advanced---such as during the Industrial Revolution---the earnings
gap between skilled and unskilled labor widens. Education begins to
offer significant advantages in the labor market. In response, parents
adjust by having fewer children and placing greater emphasis on each
child's development, including school attendance and better health care.

Evidently, industrialization raises the economic value of skilled labor,
which alters household incentives. As returns to education increase,
parents begin to favor investments in child quality over child quantity.
This shift results in declining fertility because families choose to
have fewer children and allocate more resources to each. The feedback
effect is significant: higher educational investment raises productivity
in the next generation, which in turn accelerates technological
advancement and further increases the returns to human capital. Over
time, the economy moves from a state of high fertility and low growth to
one characterized by low fertility and sustained growth.

Furthermore, \citet{galor2002natural} introduced an evolutionary
refinement to the unified growth framework by accounting for
heterogeneity in parental preferences. During the Malthusian period,
some families placed greater emphasis on child quality, such as
education, while others prioritized quantity. In a stagnant economy with
limited returns to education, high-fertility lineages maintained a
numerical advantage and suppressed average human capital. As
technological progress increased the returns to education, families that
valued quality gained an economic edge. Their children acquired more
human capital and achieved higher income and survival rates. These
advantages allowed such families to grow in relative size. Over time,
this process resembled a form of evolutionary selection, gradually
favoring quality-oriented parental types and shifting the population
toward greater emphasis on child human capital. These dynamics
strengthened the shift from high-fertility, low-education regimes to
low-fertility, high-investment family structures. In formal
overlapping-generations models, \citet{galor2002natural} demonstrate
that this evolutionary adaptation accelerates the demographic
transition. Their framework accounts for the rapid and widespread drop
in fertility once it takes effect. Higher returns to human capital push
parents to favor quality, while preferences for quality begin to
dominate within the population. These forces support the emergence of a
low-fertility, high-investment equilibrium and establish a unified
explanation for both economic development and demographic change.

Importantly, unified growth models identify several complementary
mechanisms that reinforce the basic Q--Q trade-off during development.
One is the decline of child labor. As the economy modernizes, the value
of child labor falls, both because legal reforms often restrict child
labor and because parents realize the earnings their children could make
as unskilled laborers are paltry compared to the potential returns if
those children instead spend time in school. \citet{hazan2002child}
formally show that when child labor becomes less profitable relative to
adult (skilled) labor, parents further reduce fertility and invest more
in each child's education. Historical evidence from England, for
instance, indicates that during industrialization the wages of children
(relative to adults) dropped significantly, especially in skilled
families, and this was accompanied by parents pulling children out of
work to send them to school. \citet{galor2006human} even argue that
capitalist industrialists supported public education laws and child
labor bans as a way to increase the human capital of the workforce,
inadvertently hastening the fertility transition.

Another mechanism is the rise in life expectancy and child survival.
Improvements in sanitation, nutrition, and medical knowledge in
developing societies led to more children surviving to adulthood. While
the earliest unified growth models treated mortality as exogenous or
ignored it, later research demonstrated that declining child mortality
can trigger lower fertility as well-- parents no longer need ``extra''
births for insurance once they are confident their existing children
will survive. In other words, increased child survival and the
quality--quantity trade-off are complementary explanations for fertility
decline that operate in tandem. When fewer births are lost to disease,
parents can achieve a desired number of surviving offspring with fewer
total births, and they tend to reallocate effort into each child's
health and education.

The overall effect is a reinforcing cycle: better health raises the
returns to schooling (healthier children can learn more effectively and
have longer working lives), which further encourages educational
investments and reduces fertility. Indeed, Galor notes that human
capital should be interpreted broadly to include health as well as
schooling; in unified growth theory, improvements in nutrition and
physical well-being were crucial to making labor more productive and
thus were part and parcel of the rise in demand for human capital.

The unified growth literature places the Q--Q model within a more
general account of economic and demographic change. In this framework,
higher income or stronger returns to child quality reduce fertility and
help shift economies from stagnation toward sustained growth. Several
mechanisms support this transition, such as a fall in child labor, a
drop in child mortality, and a shift in parental priorities. These
models explain not only the presence of a quantity--quality trade-off
but also its rising influence at a specific point in history. The
evidence supports these claims: countries that saw earlier increases in
returns to education experienced earlier fertility decline, while delays
in reforms, such as public education or health access, corresponded to
prolonged high fertility. As a result, the Q--Q mechanism forms a key
component of unified growth theory.

\subsubsection{Recent Refinements to the Q-Q Model}

Contemporary research has further refined the quantity--quality model by
relaxing some of its initial simplifying assumptions. Three important
extensions involve (1) capital market imperfections that constrain
parents' ability to invest in child quality, (2) intra-household
conflict and bargaining between mothers and fathers over fertility
choices, and (3) recognition that child quality is multi-dimensional,
which extends beyond schooling to include health, nutrition, and other
facets of human capital.

To begin, I examine how credit constraints can give rise to poverty
traps. The canonical quantity--quality (Q--Q) framework assumes that
parents can reallocate resources freely; borrowing against future
earnings to finance schooling or health investments whenever the
expected return is high. In practice, especially in low-income settings,
credit markets function imperfectly: poor households typically cannot
secure loans to cover children's education or medical costs even when
such investments would yield substantial future gains. This market
failure magnifies the Q--Q trade-off. Becker, Lewis, and Willis--already
noted by \citet{grawe2008quality} as emphasizing ``resource
limitations''--implicitly recognized this issue, but contemporary models
make it explicit by imposing a borrowing constraint. Parents must fund
childrearing from current income alone; they cannot collateralize a
child's future wages to pay today's school fees. Consequently, when
income is low, each additional birth directly reduces the attainable
quality per child, potentially trapping families in a
low-income--high-fertility equilibrium.

Formal models show that ``the quality--quantity trade-off arises from a
binding credit constraint that prevents parents from borrowing against
future child income.'' Empirical work supports this mechanism.
\citet{kremer2002income} and \citet{de2003inequality} document that
countries facing tighter liquidity constraints tend to display higher
fertility and lower educational attainment, consistent with
liquidity-constrained parents favoring quantity over quality.
Cross-country evidence likewise indicates that where financial frictions
are more severe, the negative correlation between fertility and
schooling is stronger. Theoretically, introducing a borrowing limit can
generate multiple steady states: one with low fertility and high
investment when incomes suffice to cover quality costs, and another with
high fertility and minimal investment when they do not. Policy
instruments such as education subsidies or conditional cash transfers
effectively relax these constraints, which nudge households toward the
low-fertility, high-investment equilibrium. In short, incorporating
credit market imperfections deepens the explanatory power of the Q-Q
model: economic growth alone may not reduce fertility if households
remain too cash-poor to afford schooling, whereas targeted
quality-enhancing transfers can catalyze both demographic and
human-capital transitions.

\bibliographystyle{aea}
\bibliography{references}

\% The appendix command is issued once, prior to all appendices, if any.
\appendix

\section{Mathematical Appendix}

\end{document}