### 2025-07-29 data-analysis/outputs/dhs_summary_birth_size.tex
add summary tables for birth size, weight, and vaccines
Added LaTeX tables summarizing perceived birth size, birth weight, and number of vaccines received by location and policy period. These tables provide descriptive statistics for further analysis and reporting.

\begin{table}[!h]
\centering
\caption{Perceived Birth Size by Location and Policy Period}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}m{3cm}cccccc}
\toprule
Variable & Manila, Pre-Policy & Manila, Post-Policy & Other NCR, Pre-Policy & Other NCR, Post-Policy & Non-NCR, Pre-Policy & Non-NCR, Post-Policy\\
\midrule
1 - Very small & 10 (4.8\%) & 1 (0.6\%) & 2 (1.6\%) & NA & 345 (4.7\%) & 116 (3.4\%)\\
2 - Smaller than average & 35 (16.8\%) & 39 (23.6\%) & 25 (20.3\%) & 34 (25.0\%) & 1256 (17.0\%) & 584 (17.2\%)\\
3 - Average size & 122 (58.7\%) & 92 (55.8\%) & 76 (61.8\%) & 84 (61.8\%) & 4397 (59.4\%) & 1990 (58.6\%)\\
4 - Larger than average & 28 (13.5\%) & 22 (13.3\%) & 18 (14.6\%) & 15 (11.0\%) & 970 (13.1\%) & 509 (15.0\%)\\
5 - Very large & 12 (5.8\%) & 11 (6.7\%) & 2 (1.6\%) & 3 (2.2\%) & 407 (5.5\%) & 184 (5.4\%)\\
\addlinespace
9 - Missing & 1 (0.5\%) & NA & NA & NA & 18 (0.2\%) & NA\\
8 - Don't know & NA & NA & NA & NA & 13 (0.2\%) & 14 (0.4\%)\\
\bottomrule
\end{tabular}
\end{table}