### 2025-07-29 data-analysis/outputs/dhs_summary_birth_size_t.tex

\begin{table}[!h]
\centering
\caption{Perceived Birth Size by Location and Policy Period}
\centering
\begin{tabular}[t]{lccccc}
\toprule
Group & \makecell{Very\\small} & \makecell{Smaller\\than\\average} & \makecell{Average\\size} & \makecell{Larger\\than\\average} & \makecell{Very\\large}\\
\midrule
Manila, Pre-Policy & \makecell{10 \\ (4.8\%)} & \makecell{35 \\ (16.9\%)} & \makecell{122 \\ (58.9\%)} & \makecell{28 \\ (13.5\%)} & \makecell{12 \\ (5.8\%)}\\
Manila, Post-Policy & \makecell{1 \\ (0.6\%)} & \makecell{39 \\ (23.6\%)} & \makecell{92 \\ (55.8\%)} & \makecell{22 \\ (13.3\%)} & \makecell{11 \\ (6.7\%)}\\
Other NCR Cities, Pre-Policy & \makecell{2 \\ (1.6\%)} & \makecell{25 \\ (20.3\%)} & \makecell{76 \\ (61.8\%)} & \makecell{18 \\ (14.6\%)} & \makecell{2 \\ (1.6\%)}\\
Other NCR Cities, Post-Policy & NA & \makecell{34 \\ (25.0\%)} & \makecell{84 \\ (61.8\%)} & \makecell{15 \\ (11.0\%)} & \makecell{3 \\ (2.2\%)}\\
Rest of the Philippines, Pre-Policy & \makecell{345 \\ (4.7\%)} & \makecell{1256 \\ (17.0\%)} & \makecell{4397 \\ (59.6\%)} & \makecell{970 \\ (13.2\%)} & \makecell{407 \\ (5.5\%)}\\
Rest of the Philippines, Post-Policy & \makecell{116 \\ (3.4\%)} & \makecell{584 \\ (17.3\%)} & \makecell{1990 \\ (58.8\%)} & \makecell{509 \\ (15.0\%)} & \makecell{184 \\ (5.4\%)}\\
\bottomrule
\end{tabular}
\end{table}