### 2025-07-29 data-analysis/outputs/dhs_summary_birth_weight_t.tex

\begin{table}[!h]
\centering
\caption{Birth Weight by Location and Policy Period}
\centering
\begin{tabular}[t]{l cl c}
\toprule
Group & Birth weight (g)\\
\midrule
Manila, Pre-Policy & 3023.12 (712.62), n = 156\\
Manila, Post-Policy & 2969.79 (619.23), n = 144\\
Other NCR Cities, Pre-Policy & 2830.41 (743.57), n = 70\\
Other NCR Cities, Post-Policy & 2965.89 (793.72), n = 88\\
Rest of the Philippines, Pre-Policy & 3106.41 (726.91), n = 4562\\
\addlinespace
Rest of the Philippines, Post-Policy & 2994.45 (681.71), n = 2298\\
\bottomrule
\end{tabular}
\end{table}