### 2025-07-29 data-analysis/outputs/dhs_summary_vaccines.tex

\begin{table}[!h]
\centering
\caption{Number of Vaccines Received by Location and Policy Period}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}m{3cm}cccccc}
\toprule
Variable & Manila, Pre-Policy & Manila, Post-Policy & Other NCR, Pre-Policy & Other NCR, Post-Policy & Non-NCR, Pre-Policy & Non-NCR, Post-Policy\\
\midrule
1–2 Vaccines & 1 (0.5\%) & 3 (1.8\%) & 1 (0.8\%) & 3 (2.2\%) & 68 (0.9\%) & 95 (2.8\%)\\
3–4 Vaccines & 4 (1.9\%) & 5 (3.0\%) & 3 (2.4\%) & 5 (3.7\%) & 136 (1.8\%) & 117 (3.4\%)\\
5+ Vaccines & 38 (18.2\%) & 33 (20.0\%) & 31 (25.2\%) & 52 (38.2\%) & 1869 (25.1\%) & 1141 (33.5\%)\\
No Vaccines & 166 (79.4\%) & 124 (75.2\%) & 88 (71.5\%) & 76 (55.9\%) & 5364 (72.1\%) & 2057 (60.3\%)\\
\bottomrule
\end{tabular}
\end{table}