### 2025-07-29 data-analysis/outputs/dhs_summary_vaccines_t.tex

\begin{table}[!h]
\centering
\caption{Number of Vaccines Received by Location and Policy Period}
\centering
\begin{tabular}[t]{lcccc}
\toprule
Group & 1–2 Vaccines & 3–4 Vaccines & 5+ Vaccines & No Vaccines\\
\midrule
Manila, Pre-Policy & 1 (0.5\%) & 4 (1.9\%) & 38 (18.2\%) & 166 (79.4\%)\\
Manila, Post-Policy & 3 (1.8\%) & 5 (3.0\%) & 33 (20.0\%) & 124 (75.2\%)\\
Other NCR Cities, Pre-Policy & 1 (0.8\%) & 3 (2.4\%) & 31 (25.2\%) & 88 (71.5\%)\\
Other NCR Cities, Post-Policy & 3 (2.2\%) & 5 (3.7\%) & 52 (38.2\%) & 76 (55.9\%)\\
Rest of the Philippines, Pre-Policy & 68 (0.9\%) & 136 (1.8\%) & 1869 (25.1\%) & 5364 (72.1\%)\\
\addlinespace
Rest of the Philippines, Post-Policy & 95 (2.8\%) & 117 (3.4\%) & 1141 (33.5\%) & 2057 (60.3\%)\\
\bottomrule
\end{tabular}
\end{table}