### 2025-08-07 data-analysis/outputs/dhs_descriptive_combined.tex
add output tables for birthrate and DHS descriptive stats
Added LaTeX output files for census birthrate changes and various DHS descriptive and summary statistics tables. These outputs summarize birth rates by region and period, and provide detailed descriptive statistics for outcomes by location and policy period.

\begin{table}[!h]
\centering
\caption{Descriptive statistics by location and policy period (outcomes only).}
\centering
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{lcccccc}
\toprule
\multicolumn{1}{c}{\textbf{ }} & \multicolumn{151}{c}{\textbf{Manila, Post-Policy}} & \multicolumn{209}{c}{\textbf{Manila, Pre-Policy}} & \multicolumn{3410}{c}{\textbf{Non-NCR, Post-Policy}} & \multicolumn{7437}{c}{\textbf{Non-NCR, Pre-Policy}} & \multicolumn{136}{c}{\textbf{Other NCR, Post-Policy}} & \multicolumn{123}{c}{\textbf{Other NCR, Pre-Policy}} & \multicolumn{10334}{c}{\textbf{NA}} \\
\cmidrule(l{3pt}r{3pt}){2-152} \cmidrule(l{3pt}r{3pt}){153-361} \cmidrule(l{3pt}r{3pt}){362-3771} \cmidrule(l{3pt}r{3pt}){3772-11208} \cmidrule(l{3pt}r{3pt}){11209-11344} \cmidrule(l{3pt}r{3pt}){11345-11467} NA
Variable & 123 & 136 & 151 & 209 & 3410 & 7437\\
\midrule
\cellcolor{gray!10}{Birth weight (grams)} & \cellcolor{gray!10}{5816.11 \\ (3592.43)} & \cellcolor{gray!10}{5379.28 \\ (3411.51)} & \cellcolor{gray!10}{3937.93 \\ (2439.54)} & \cellcolor{gray!10}{4766.51 \\ (3089.10)} & \cellcolor{gray!10}{5254.08 \\ (3321.64)} & \cellcolor{gray!10}{5754.47 \\ (3399.87)}\\
1 - Very small & 2 (1.6%) & NA & 1 (0.7%) & 10 (4.8%) & 116 (3.4%) & 345 (4.7%)\\
\cellcolor{gray!10}{2 - Smaller than average} & \cellcolor{gray!10}{25 (20.3%)} & \cellcolor{gray!10}{34 (25.0%)} & \cellcolor{gray!10}{38 (25.2%)} & \cellcolor{gray!10}{35 (16.8%)} & \cellcolor{gray!10}{584 (17.2%)} & \cellcolor{gray!10}{1256 (17.0%)}\\
3 - Average size & 76 (61.8%) & 84 (61.8%) & 84 (55.6%) & 122 (58.7%) & 1990 (58.6%) & 4397 (59.4%)\\
\cellcolor{gray!10}{4 - Larger than average} & \cellcolor{gray!10}{18 (14.6%)} & \cellcolor{gray!10}{15 (11.0%)} & \cellcolor{gray!10}{18 (11.9%)} & \cellcolor{gray!10}{28 (13.5%)} & \cellcolor{gray!10}{509 (15.0%)} & \cellcolor{gray!10}{970 (13.1%)}\\
5 - Very large & 2 (1.6%) & 3 (2.2%) & 10 (6.6%) & 12 (5.8%) & 184 (5.4%) & 407 (5.5%)\\
\cellcolor{gray!10}{9 - Missing} & \cellcolor{gray!10}{NA} & \cellcolor{gray!10}{NA} & \cellcolor{gray!10}{NA} & \cellcolor{gray!10}{1 (0.5%)} & \cellcolor{gray!10}{NA} & \cellcolor{gray!10}{18 (0.2%)}\\
8 - Don’t know & NA & NA & NA & NA & 14 (0.4%) & 13 (0.2%)\\
\cellcolor{gray!10}{1–2 Vaccines} & \cellcolor{gray!10}{1 (0.8%)} & \cellcolor{gray!10}{3 (2.2%)} & \cellcolor{gray!10}{3 (2.0%)} & \cellcolor{gray!10}{1 (0.5%)} & \cellcolor{gray!10}{95 (2.8%)} & \cellcolor{gray!10}{68 (0.9%)}\\
3–4 Vaccines & 3 (2.4%) & 5 (3.7%) & 5 (3.3%) & 4 (1.9%) & 117 (3.4%) & 136 (1.8%)\\
\cellcolor{gray!10}{5+ Vaccines} & \cellcolor{gray!10}{31 (25.2%)} & \cellcolor{gray!10}{52 (38.2%)} & \cellcolor{gray!10}{29 (19.2%)} & \cellcolor{gray!10}{38 (18.2%)} & \cellcolor{gray!10}{1141 (33.5%)} & \cellcolor{gray!10}{1869 (25.1%)}\\
No Vaccines & 88 (71.5%) & 76 (55.9%) & 114 (75.5%) & 166 (79.4%) & 2057 (60.3%) & 5364 (72.1%)\\
\bottomrule
\end{tabular}
\end{table}