### 2025-08-07 data-analysis/outputs/dhs_summary_combined_table.tex

\begin{table}[!h]
\centering
\caption{Descriptive statistics by location and policy period (outcomes only).}
\centering
\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
\fontsize{8}{10}\selectfont
\begin{tabular}[t]{llrrlllllllrrllll}
\toprule
Group & Birth weight (grams) & n.x & pct.x & 1 - Very small & 2 - Smaller than average & 3 - Average size & 4 - Larger than average & 5 - Very large & 9 - Missing & 8 - Don't know & n.y & pct.y & 1–2 Vaccines & 3–4 Vaccines & 5+ Vaccines & No Vaccines\\
\midrule
Manila, Post-Policy & 3864 (2419) & 1 & 0.0060606 & 1 (0.6%) & NA & NA & NA & NA & NA & NA & 3 & 0.0181818 & 3 (1.8%) & NA & NA & NA\\
Manila, Post-Policy & 3864 (2419) & 1 & 0.0060606 & 1 (0.6%) & NA & NA & NA & NA & NA & NA & 5 & 0.0303030 & NA & 5 (3.0%) & NA & NA\\
Manila, Post-Policy & 3864 (2419) & 1 & 0.0060606 & 1 (0.6%) & NA & NA & NA & NA & NA & NA & 33 & 0.2000000 & NA & NA & 33 (20.0%) & NA\\
Manila, Post-Policy & 3864 (2419) & 1 & 0.0060606 & 1 (0.6%) & NA & NA & NA & NA & NA & NA & 124 & 0.7515152 & NA & NA & NA & 124 (75.2%)\\
Manila, Post-Policy & 3864 (2419) & 39 & 0.2363636 & NA & 39 (23.6%) & NA & NA & NA & NA & NA & 3 & 0.0181818 & 3 (1.8%) & NA & NA & NA\\
Manila, Post-Policy & 3864 (2419) & 39 & 0.2363636 & NA & 39 (23.6%) & NA & NA & NA & NA & NA & 5 & 0.0303030 & NA & 5 (3.0%) & NA & NA\\
Manila, Post-Policy & 3864 (2419) & 39 & 0.2363636 & NA & 39 (23.6%) & NA & NA & NA & NA & NA & 33 & 0.2000000 & NA & NA & 33 (20.0%) & NA\\
Manila, Post-Policy & 3864 (2419) & 39 & 0.2363636 & NA & 39 (23.6%) & NA & NA & NA & NA & NA & 124 & 0.7515152 & NA & NA & NA & 124 (75.2%)\\
Manila, Post-Policy & 3864 (2419) & 92 & 0.5575758 & NA & NA & 92 (55.8%) & NA & NA & NA & NA & 3 & 0.0181818 & 3 (1.8%) & NA & NA & NA\\
Manila, Post-Policy & 3864 (2419) & 92 & 0.5575758 & NA & NA & 92 (55.8%) & NA & NA & NA & NA & 5 & 0.0303030 & NA & 5 (3.0%) & NA & NA\\
Manila, Post-Policy & 3864 (2419) & 92 & 0.5575758 & NA & NA & 92 (55.8%) & NA & NA & NA & NA & 33 & 0.2000000 & NA & NA & 33 (20.0%) & NA\\
Manila, Post-Policy & 3864 (2419) & 92 & 0.5575758 & NA & NA & 92 (55.8%) & NA & NA & NA & NA & 124 & 0.7515152 & NA & NA & NA & 124 (75.2%)\\
Manila, Post-Policy & 3864 (2419) & 22 & 0.1333333 & NA & NA & NA & 22 (13.3%) & NA & NA & NA & 3 & 0.0181818 & 3 (1.8%) & NA & NA & NA\\
Manila, Post-Policy & 3864 (2419) & 22 & 0.1333333 & NA & NA & NA & 22 (13.3%) & NA & NA & NA & 5 & 0.0303030 & NA & 5 (3.0%) & NA & NA\\
Manila, Post-Policy & 3864 (2419) & 22 & 0.1333333 & NA & NA & NA & 22 (13.3%) & NA & NA & NA & 33 & 0.2000000 & NA & NA & 33 (20.0%) & NA\\
Manila, Post-Policy & 3864 (2419) & 22 & 0.1333333 & NA & NA & NA & 22 (13.3%) & NA & NA & NA & 124 & 0.7515152 & NA & NA & NA & 124 (75.2%)\\
Manila, Post-Policy & 3864 (2419) & 11 & 0.0666667 & NA & NA & NA & NA & 11 (6.7%) & NA & NA & 3 & 0.0181818 & 3 (1.8%) & NA & NA & NA\\
Manila, Post-Policy & 3864 (2419) & 11 & 0.0666667 & NA & NA & NA & NA & 11 (6.7%) & NA & NA & 5 & 0.0303030 & NA & 5 (3.0%) & NA & NA\\
Manila, Post-Policy & 3864 (2419) & 11 & 0.0666667 & NA & NA & NA & NA & 11 (6.7%) & NA & NA & 33 & 0.2000000 & NA & NA & 33 (20.0%) & NA\\
Manila, Post-Policy & 3864 (2419) & 11 & 0.0666667 & NA & NA & NA & NA & 11 (6.7%) & NA & NA & 124 & 0.7515152 & NA & NA & NA & 124 (75.2%)\\
Manila, Pre-Policy & 4767 (3089) & 10 & 0.0480769 & 10 (4.8%) & NA & NA & NA & NA & NA & NA & 1 & 0.0047847 & 1 (0.5%) & NA & NA & NA\\
Manila, Pre-Policy & 4767 (3089) & 10 & 0.0480769 & 10 (4.8%) & NA & NA & NA & NA & NA & NA & 4 & 0.0191388 & NA & 4 (1.9%) & NA & NA\\
Manila, Pre-Policy & 4767 (3089) & 10 & 0.0480769 & 10 (4.8%) & NA & NA & NA & NA & NA & NA & 38 & 0.1818182 & NA & NA & 38 (18.2%) & NA\\
Manila, Pre-Policy & 4767 (3089) & 10 & 0.0480769 & 10 (4.8%) & NA & NA & NA & NA & NA & NA & 166 & 0.7942584 & NA & NA & NA & 166 (79.4%)\\
Manila, Pre-Policy & 4767 (3089) & 35 & 0.1682692 & NA & 35 (16.8%) & NA & NA & NA & NA & NA & 1 & 0.0047847 & 1 (0.5%) & NA & NA & NA\\
Manila, Pre-Policy & 4767 (3089) & 35 & 0.1682692 & NA & 35 (16.8%) & NA & NA & NA & NA & NA & 4 & 0.0191388 & NA & 4 (1.9%) & NA & NA\\
Manila, Pre-Policy & 4767 (3089) & 35 & 0.1682692 & NA & 35 (16.8%) & NA & NA & NA & NA & NA & 38 & 0.1818182 & NA & NA & 38 (18.2%) & NA\\
Manila, Pre-Policy & 4767 (3089) & 35 & 0.1682692 & NA & 35 (16.8%) & NA & NA & NA & NA & NA & 166 & 0.7942584 & NA & NA & NA & 166 (79.4%)\\
Manila, Pre-Policy & 4767 (3089) & 122 & 0.5865385 & NA & NA & 122 (58.7%) & NA & NA & NA & NA & 1 & 0.0047847 & 1 (0.5%) & NA & NA & NA\\
Manila, Pre-Policy & 4767 (3089) & 122 & 0.5865385 & NA & NA & 122 (58.7%) & NA & NA & NA & NA & 4 & 0.0191388 & NA & 4 (1.9%) & NA & NA\\
Manila, Pre-Policy & 4767 (3089) & 122 & 0.5865385 & NA & NA & 122 (58.7%) & NA & NA & NA & NA & 38 & 0.1818182 & NA & NA & 38 (18.2%) & NA\\
Manila, Pre-Policy & 4767 (3089) & 122 & 0.5865385 & NA & NA & 122 (58.7%) & NA & NA & NA & NA & 166 & 0.7942584 & NA & NA & NA & 166 (79.4%)\\
Manila, Pre-Policy & 4767 (3089) & 28 & 0.1346154 & NA & NA & NA & 28 (13.5%) & NA & NA & NA & 1 & 0.0047847 & 1 (0.5%) & NA & NA & NA\\
Manila, Pre-Policy & 4767 (3089) & 28 & 0.1346154 & NA & NA & NA & 28 (13.5%) & NA & NA & NA & 4 & 0.0191388 & NA & 4 (1.9%) & NA & NA\\
Manila, Pre-Policy & 4767 (3089) & 28 & 0.1346154 & NA & NA & NA & 28 (13.5%) & NA & NA & NA & 38 & 0.1818182 & NA & NA & 38 (18.2%) & NA\\
Manila, Pre-Policy & 4767 (3089) & 28 & 0.1346154 & NA & NA & NA & 28 (13.5%) & NA & NA & NA & 166 & 0.7942584 & NA & NA & NA & 166 (79.4%)\\
Manila, Pre-Policy & 4767 (3089) & 12 & 0.0576923 & NA & NA & NA & NA & 12 (5.8%) & NA & NA & 1 & 0.0047847 & 1 (0.5%) & NA & NA & NA\\
Manila, Pre-Policy & 4767 (3089) & 12 & 0.0576923 & NA & NA & NA & NA & 12 (5.8%) & NA & NA & 4 & 0.0191388 & NA & 4 (1.9%) & NA & NA\\
Manila, Pre-Policy & 4767 (3089) & 12 & 0.0576923 & NA & NA & NA & NA & 12 (5.8%) & NA & NA & 38 & 0.1818182 & NA & NA & 38 (18.2%) & NA\\
Manila, Pre-Policy & 4767 (3089) & 12 & 0.0576923 & NA & NA & NA & NA & 12 (5.8%) & NA & NA & 166 & 0.7942584 & NA & NA & NA & 166 (79.4%)\\
Manila, Pre-Policy & 4767 (3089) & 1 & 0.0048077 & NA & NA & NA & NA & NA & 1 (0.5%) & NA & 1 & 0.0047847 & 1 (0.5%) & NA & NA & NA\\
Manila, Pre-Policy & 4767 (3089) & 1 & 0.0048077 & NA & NA & NA & NA & NA & 1 (0.5%) & NA & 4 & 0.0191388 & NA & 4 (1.9%) & NA & NA\\
Manila, Pre-Policy & 4767 (3089) & 1 & 0.0048077 & NA & NA & NA & NA & NA & 1 (0.5%) & NA & 38 & 0.1818182 & NA & NA & 38 (18.2%) & NA\\
Manila, Pre-Policy & 4767 (3089) & 1 & 0.0048077 & NA & NA & NA & NA & NA & 1 (0.5%) & NA & 166 & 0.7942584 & NA & NA & NA & 166 (79.4%)\\
Non-NCR, Post-Policy & 5254 (3322) & 116 & 0.0341478 & 116 (3.4%) & NA & NA & NA & NA & NA & NA & 95 & 0.0278592 & 95 (2.8%) & NA & NA & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 116 & 0.0341478 & 116 (3.4%) & NA & NA & NA & NA & NA & NA & 117 & 0.0343109 & NA & 117 (3.4%) & NA & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 116 & 0.0341478 & 116 (3.4%) & NA & NA & NA & NA & NA & NA & 1141 & 0.3346041 & NA & NA & 1141 (33.5%) & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 116 & 0.0341478 & 116 (3.4%) & NA & NA & NA & NA & NA & NA & 2057 & 0.6032258 & NA & NA & NA & 2057 (60.3%)\\
Non-NCR, Post-Policy & 5254 (3322) & 584 & 0.1719164 & NA & 584 (17.2%) & NA & NA & NA & NA & NA & 95 & 0.0278592 & 95 (2.8%) & NA & NA & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 584 & 0.1719164 & NA & 584 (17.2%) & NA & NA & NA & NA & NA & 117 & 0.0343109 & NA & 117 (3.4%) & NA & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 584 & 0.1719164 & NA & 584 (17.2%) & NA & NA & NA & NA & NA & 1141 & 0.3346041 & NA & NA & 1141 (33.5%) & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 584 & 0.1719164 & NA & 584 (17.2%) & NA & NA & NA & NA & NA & 2057 & 0.6032258 & NA & NA & NA & 2057 (60.3%)\\
Non-NCR, Post-Policy & 5254 (3322) & 1990 & 0.5858110 & NA & NA & 1990 (58.6%) & NA & NA & NA & NA & 95 & 0.0278592 & 95 (2.8%) & NA & NA & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 1990 & 0.5858110 & NA & NA & 1990 (58.6%) & NA & NA & NA & NA & 117 & 0.0343109 & NA & 117 (3.4%) & NA & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 1990 & 0.5858110 & NA & NA & 1990 (58.6%) & NA & NA & NA & NA & 1141 & 0.3346041 & NA & NA & 1141 (33.5%) & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 1990 & 0.5858110 & NA & NA & 1990 (58.6%) & NA & NA & NA & NA & 2057 & 0.6032258 & NA & NA & NA & 2057 (60.3%)\\
Non-NCR, Post-Policy & 5254 (3322) & 509 & 0.1498381 & NA & NA & NA & 509 (15.0%) & NA & NA & NA & 95 & 0.0278592 & 95 (2.8%) & NA & NA & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 509 & 0.1498381 & NA & NA & NA & 509 (15.0%) & NA & NA & NA & 117 & 0.0343109 & NA & 117 (3.4%) & NA & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 509 & 0.1498381 & NA & NA & NA & 509 (15.0%) & NA & NA & NA & 1141 & 0.3346041 & NA & NA & 1141 (33.5%) & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 509 & 0.1498381 & NA & NA & NA & 509 (15.0%) & NA & NA & NA & 2057 & 0.6032258 & NA & NA & NA & 2057 (60.3%)\\
Non-NCR, Post-Policy & 5254 (3322) & 184 & 0.0541654 & NA & NA & NA & NA & 184 (5.4%) & NA & NA & 95 & 0.0278592 & 95 (2.8%) & NA & NA & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 184 & 0.0541654 & NA & NA & NA & NA & 184 (5.4%) & NA & NA & 117 & 0.0343109 & NA & 117 (3.4%) & NA & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 184 & 0.0541654 & NA & NA & NA & NA & 184 (5.4%) & NA & NA & 1141 & 0.3346041 & NA & NA & 1141 (33.5%) & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 184 & 0.0541654 & NA & NA & NA & NA & 184 (5.4%) & NA & NA & 2057 & 0.6032258 & NA & NA & NA & 2057 (60.3%)\\
Non-NCR, Post-Policy & 5254 (3322) & 14 & 0.0041213 & NA & NA & NA & NA & NA & NA & 14 (0.4%) & 95 & 0.0278592 & 95 (2.8%) & NA & NA & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 14 & 0.0041213 & NA & NA & NA & NA & NA & NA & 14 (0.4%) & 117 & 0.0343109 & NA & 117 (3.4%) & NA & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 14 & 0.0041213 & NA & NA & NA & NA & NA & NA & 14 (0.4%) & 1141 & 0.3346041 & NA & NA & 1141 (33.5%) & NA\\
Non-NCR, Post-Policy & 5254 (3322) & 14 & 0.0041213 & NA & NA & NA & NA & NA & NA & 14 (0.4%) & 2057 & 0.6032258 & NA & NA & NA & 2057 (60.3%)\\
Non-NCR, Pre-Policy & 5754 (3400) & 345 & 0.0465839 & 345 (4.7%) & NA & NA & NA & NA & NA & NA & 68 & 0.0091435 & 68 (0.9%) & NA & NA & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 345 & 0.0465839 & 345 (4.7%) & NA & NA & NA & NA & NA & NA & 136 & 0.0182869 & NA & 136 (1.8%) & NA & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 345 & 0.0465839 & 345 (4.7%) & NA & NA & NA & NA & NA & NA & 1869 & 0.2513110 & NA & NA & 1869 (25.1%) & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 345 & 0.0465839 & 345 (4.7%) & NA & NA & NA & NA & NA & NA & 5364 & 0.7212586 & NA & NA & NA & 5364 (72.1%)\\
Non-NCR, Pre-Policy & 5754 (3400) & 1256 & 0.1695922 & NA & 1256 (17.0%) & NA & NA & NA & NA & NA & 68 & 0.0091435 & 68 (0.9%) & NA & NA & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 1256 & 0.1695922 & NA & 1256 (17.0%) & NA & NA & NA & NA & NA & 136 & 0.0182869 & NA & 136 (1.8%) & NA & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 1256 & 0.1695922 & NA & 1256 (17.0%) & NA & NA & NA & NA & NA & 1869 & 0.2513110 & NA & NA & 1869 (25.1%) & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 1256 & 0.1695922 & NA & 1256 (17.0%) & NA & NA & NA & NA & NA & 5364 & 0.7212586 & NA & NA & NA & 5364 (72.1%)\\
Non-NCR, Pre-Policy & 5754 (3400) & 4397 & 0.5937078 & NA & NA & 4397 (59.4%) & NA & NA & NA & NA & 68 & 0.0091435 & 68 (0.9%) & NA & NA & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 4397 & 0.5937078 & NA & NA & 4397 (59.4%) & NA & NA & NA & NA & 136 & 0.0182869 & NA & 136 (1.8%) & NA & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 4397 & 0.5937078 & NA & NA & 4397 (59.4%) & NA & NA & NA & NA & 1869 & 0.2513110 & NA & NA & 1869 (25.1%) & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 4397 & 0.5937078 & NA & NA & 4397 (59.4%) & NA & NA & NA & NA & 5364 & 0.7212586 & NA & NA & NA & 5364 (72.1%)\\
Non-NCR, Pre-Policy & 5754 (3400) & 970 & 0.1309749 & NA & NA & NA & 970 (13.1%) & NA & NA & NA & 68 & 0.0091435 & 68 (0.9%) & NA & NA & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 970 & 0.1309749 & NA & NA & NA & 970 (13.1%) & NA & NA & NA & 136 & 0.0182869 & NA & 136 (1.8%) & NA & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 970 & 0.1309749 & NA & NA & NA & 970 (13.1%) & NA & NA & NA & 1869 & 0.2513110 & NA & NA & 1869 (25.1%) & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 970 & 0.1309749 & NA & NA & NA & 970 (13.1%) & NA & NA & NA & 5364 & 0.7212586 & NA & NA & NA & 5364 (72.1%)\\
Non-NCR, Pre-Policy & 5754 (3400) & 407 & 0.0549554 & NA & NA & NA & NA & 407 (5.5%) & NA & NA & 68 & 0.0091435 & 68 (0.9%) & NA & NA & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 407 & 0.0549554 & NA & NA & NA & NA & 407 (5.5%) & NA & NA & 136 & 0.0182869 & NA & 136 (1.8%) & NA & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 407 & 0.0549554 & NA & NA & NA & NA & 407 (5.5%) & NA & NA & 1869 & 0.2513110 & NA & NA & 1869 (25.1%) & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 407 & 0.0549554 & NA & NA & NA & NA & 407 (5.5%) & NA & NA & 5364 & 0.7212586 & NA & NA & NA & 5364 (72.1%)\\
Non-NCR, Pre-Policy & 5754 (3400) & 13 & 0.0017553 & NA & NA & NA & NA & NA & NA & 13 (0.2%) & 68 & 0.0091435 & 68 (0.9%) & NA & NA & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 13 & 0.0017553 & NA & NA & NA & NA & NA & NA & 13 (0.2%) & 136 & 0.0182869 & NA & 136 (1.8%) & NA & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 13 & 0.0017553 & NA & NA & NA & NA & NA & NA & 13 (0.2%) & 1869 & 0.2513110 & NA & NA & 1869 (25.1%) & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 13 & 0.0017553 & NA & NA & NA & NA & NA & NA & 13 (0.2%) & 5364 & 0.7212586 & NA & NA & NA & 5364 (72.1%)\\
Non-NCR, Pre-Policy & 5754 (3400) & 18 & 0.0024305 & NA & NA & NA & NA & NA & 18 (0.2%) & NA & 68 & 0.0091435 & 68 (0.9%) & NA & NA & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 18 & 0.0024305 & NA & NA & NA & NA & NA & 18 (0.2%) & NA & 136 & 0.0182869 & NA & 136 (1.8%) & NA & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 18 & 0.0024305 & NA & NA & NA & NA & NA & 18 (0.2%) & NA & 1869 & 0.2513110 & NA & NA & 1869 (25.1%) & NA\\
Non-NCR, Pre-Policy & 5754 (3400) & 18 & 0.0024305 & NA & NA & NA & NA & NA & 18 (0.2%) & NA & 5364 & 0.7212586 & NA & NA & NA & 5364 (72.1%)\\
Other NCR, Post-Policy & 5379 (3412) & 34 & 0.2500000 & NA & 34 (25.0%) & NA & NA & NA & NA & NA & 3 & 0.0220588 & 3 (2.2%) & NA & NA & NA\\
Other NCR, Post-Policy & 5379 (3412) & 34 & 0.2500000 & NA & 34 (25.0%) & NA & NA & NA & NA & NA & 5 & 0.0367647 & NA & 5 (3.7%) & NA & NA\\
Other NCR, Post-Policy & 5379 (3412) & 34 & 0.2500000 & NA & 34 (25.0%) & NA & NA & NA & NA & NA & 52 & 0.3823529 & NA & NA & 52 (38.2%) & NA\\
Other NCR, Post-Policy & 5379 (3412) & 34 & 0.2500000 & NA & 34 (25.0%) & NA & NA & NA & NA & NA & 76 & 0.5588235 & NA & NA & NA & 76 (55.9%)\\
Other NCR, Post-Policy & 5379 (3412) & 84 & 0.6176471 & NA & NA & 84 (61.8%) & NA & NA & NA & NA & 3 & 0.0220588 & 3 (2.2%) & NA & NA & NA\\
Other NCR, Post-Policy & 5379 (3412) & 84 & 0.6176471 & NA & NA & 84 (61.8%) & NA & NA & NA & NA & 5 & 0.0367647 & NA & 5 (3.7%) & NA & NA\\
Other NCR, Post-Policy & 5379 (3412) & 84 & 0.6176471 & NA & NA & 84 (61.8%) & NA & NA & NA & NA & 52 & 0.3823529 & NA & NA & 52 (38.2%) & NA\\
Other NCR, Post-Policy & 5379 (3412) & 84 & 0.6176471 & NA & NA & 84 (61.8%) & NA & NA & NA & NA & 76 & 0.5588235 & NA & NA & NA & 76 (55.9%)\\
Other NCR, Post-Policy & 5379 (3412) & 15 & 0.1102941 & NA & NA & NA & 15 (11.0%) & NA & NA & NA & 3 & 0.0220588 & 3 (2.2%) & NA & NA & NA\\
Other NCR, Post-Policy & 5379 (3412) & 15 & 0.1102941 & NA & NA & NA & 15 (11.0%) & NA & NA & NA & 5 & 0.0367647 & NA & 5 (3.7%) & NA & NA\\
Other NCR, Post-Policy & 5379 (3412) & 15 & 0.1102941 & NA & NA & NA & 15 (11.0%) & NA & NA & NA & 52 & 0.3823529 & NA & NA & 52 (38.2%) & NA\\
Other NCR, Post-Policy & 5379 (3412) & 15 & 0.1102941 & NA & NA & NA & 15 (11.0%) & NA & NA & NA & 76 & 0.5588235 & NA & NA & NA & 76 (55.9%)\\
Other NCR, Post-Policy & 5379 (3412) & 3 & 0.0220588 & NA & NA & NA & NA & 3 (2.2%) & NA & NA & 3 & 0.0220588 & 3 (2.2%) & NA & NA & NA\\
Other NCR, Post-Policy & 5379 (3412) & 3 & 0.0220588 & NA & NA & NA & NA & 3 (2.2%) & NA & NA & 5 & 0.0367647 & NA & 5 (3.7%) & NA & NA\\
Other NCR, Post-Policy & 5379 (3412) & 3 & 0.0220588 & NA & NA & NA & NA & 3 (2.2%) & NA & NA & 52 & 0.3823529 & NA & NA & 52 (38.2%) & NA\\
Other NCR, Post-Policy & 5379 (3412) & 3 & 0.0220588 & NA & NA & NA & NA & 3 (2.2%) & NA & NA & 76 & 0.5588235 & NA & NA & NA & 76 (55.9%)\\
Other NCR, Pre-Policy & 5816 (3592) & 2 & 0.0162602 & 2 (1.6%) & NA & NA & NA & 2 (1.6%) & NA & NA & 1 & 0.0081301 & 1 (0.8%) & NA & NA & NA\\
Other NCR, Pre-Policy & 5816 (3592) & 2 & 0.0162602 & 2 (1.6%) & NA & NA & NA & 2 (1.6%) & NA & NA & 3 & 0.0243902 & NA & 3 (2.4%) & NA & NA\\
Other NCR, Pre-Policy & 5816 (3592) & 2 & 0.0162602 & 2 (1.6%) & NA & NA & NA & 2 (1.6%) & NA & NA & 31 & 0.2520325 & NA & NA & 31 (25.2%) & NA\\
Other NCR, Pre-Policy & 5816 (3592) & 2 & 0.0162602 & 2 (1.6%) & NA & NA & NA & 2 (1.6%) & NA & NA & 88 & 0.7154472 & NA & NA & NA & 88 (71.5%)\\
Other NCR, Pre-Policy & 5816 (3592) & 25 & 0.2032520 & NA & 25 (20.3%) & NA & NA & NA & NA & NA & 1 & 0.0081301 & 1 (0.8%) & NA & NA & NA\\
Other NCR, Pre-Policy & 5816 (3592) & 25 & 0.2032520 & NA & 25 (20.3%) & NA & NA & NA & NA & NA & 3 & 0.0243902 & NA & 3 (2.4%) & NA & NA\\
Other NCR, Pre-Policy & 5816 (3592) & 25 & 0.2032520 & NA & 25 (20.3%) & NA & NA & NA & NA & NA & 31 & 0.2520325 & NA & NA & 31 (25.2%) & NA\\
Other NCR, Pre-Policy & 5816 (3592) & 25 & 0.2032520 & NA & 25 (20.3%) & NA & NA & NA & NA & NA & 88 & 0.7154472 & NA & NA & NA & 88 (71.5%)\\
Other NCR, Pre-Policy & 5816 (3592) & 76 & 0.6178862 & NA & NA & 76 (61.8%) & NA & NA & NA & NA & 1 & 0.0081301 & 1 (0.8%) & NA & NA & NA\\
Other NCR, Pre-Policy & 5816 (3592) & 76 & 0.6178862 & NA & NA & 76 (61.8%) & NA & NA & NA & NA & 3 & 0.0243902 & NA & 3 (2.4%) & NA & NA\\
Other NCR, Pre-Policy & 5816 (3592) & 76 & 0.6178862 & NA & NA & 76 (61.8%) & NA & NA & NA & NA & 31 & 0.2520325 & NA & NA & 31 (25.2%) & NA\\
Other NCR, Pre-Policy & 5816 (3592) & 76 & 0.6178862 & NA & NA & 76 (61.8%) & NA & NA & NA & NA & 88 & 0.7154472 & NA & NA & NA & 88 (71.5%)\\
Other NCR, Pre-Policy & 5816 (3592) & 18 & 0.1463415 & NA & NA & NA & 18 (14.6%) & NA & NA & NA & 1 & 0.0081301 & 1 (0.8%) & NA & NA & NA\\
Other NCR, Pre-Policy & 5816 (3592) & 18 & 0.1463415 & NA & NA & NA & 18 (14.6%) & NA & NA & NA & 3 & 0.0243902 & NA & 3 (2.4%) & NA & NA\\
Other NCR, Pre-Policy & 5816 (3592) & 18 & 0.1463415 & NA & NA & NA & 18 (14.6%) & NA & NA & NA & 31 & 0.2520325 & NA & NA & 31 (25.2%) & NA\\
Other NCR, Pre-Policy & 5816 (3592) & 18 & 0.1463415 & NA & NA & NA & 18 (14.6%) & NA & NA & NA & 88 & 0.7154472 & NA & NA & NA & 88 (71.5%)\\
\bottomrule
\end{tabular}}
\end{table}