### 2025-08-07 data-analysis/outputs/dhs_summary_table_simple.tex

[1] "Non-NCR, Pre-Policy"  "Non-NCR, Post-Policy"
[1] "Non-NCR, Pre-Policy"  "Non-NCR, Post-Policy"
  [1] "case_id"                       "cluster_id"                    "household_id"                 
  [4] "sample_weight"                 "mother_age_current"            "residence_type"               
  [7] "household_size"                "household_u5_count"            "region"                       
 [10] "province"                      "municipality"                  "wealth_index"                 
 [13] "wealth_score"                  "birth_index"                   "birth_order"                  
 [16] "is_twin"                       "birth_month"                   "birth_year"                   
 [19] "birth_date"                    "child_sex"                     "child_alive"                  
 [22] "age_at_death"                  "child_age"                     "preceding_birth_interval"     
 [25] "succeeding_birth_interval"     "delivery_place"                "birth_size"                   
 [28] "birth_weight"                  "received_bcg"                  "bcg_day"                      
 [31] "bcg_month"                     "bcg_year"                      "received_dpt1"                
 [34] "dpt1_day"                      "dpt1_month"                    "dpt1_year"                    
 [37] "received_polio1"               "polio1_day"                    "polio1_month"                 
 [40] "polio1_year"                   "received_dpt2"                 "dpt2_day"                     
 [43] "dpt2_month"                    "dpt2_year"                     "received_polio2"              
 [46] "polio2_day"                    "polio2_month"                  "polio2_year"                  
 [49] "received_dpt3"                 "dpt3_day"                      "dpt3_month"                   
 [52] "dpt3_year"                     "received_polio3"               "polio3_day"                   
 [55] "polio3_month"                  "polio3_year"                   "received_measles"             
 [58] "measles_day"                   "measles_month"                 "measles_year"                 
 [61] "ever_vaccinated"               "diarrhea_recent"               "fever_recent"                 
 [64] "cough_recent"                  "total_children_ever"           "sons_at_home"                 
 [67] "daughters_at_home"             "sons_died"                     "daughters_died"               
 [70] "births_last5"                  "births_last1"                  "births_this_month"            
 [73] "mother_age_first_birth"        "currently_pregnant"            "living_children"              
 [76] "living_plus_preg"              "living_plus_preg6"             "contraceptive_method"         
 [79] "contraceptive_intent"          "wanted_last_child"             "mother_educ_level"            
 [82] "mother_literate"               "mother_educ_attain"            "marital_status"               
 [85] "mother_age_first_sex"          "ideal_children"                "partner_educ_level"           
 [88] "partner_occupation"            "mother_occupation"             "mother_work_location"         
 [91] "survey_year"                   "region_dhs.x"                  "admin_region_name.x"          
 [94] "province_dhs.x"                "province_name.x"               "municipality_dhs.x"           
 [97] "municipality_name.x"           "barangay_name.x"               "latitude.x"                   
[100] "longitude.x"                   "altitude_gps.x"                "altitude_dem.x"               
[103] "cc_fips.x"                     "source.x"                      "urban_rural.x"                
[106] "datum.x"                       "region_dhs.y"                  "admin_region_name.y"          
[109] "province_dhs.y"                "province_name.y"               "municipality_dhs.y"           
[112] "municipality_name.y"           "barangay_name.y"               "latitude.y"                   
[115] "longitude.y"                   "altitude_gps.y"                "altitude_dem.y"               
[118] "cc_fips.y"                     "source.y"                      "urban_rural.y"                
[121] "datum.y"                       "region_dhs"                    "admin_region_name"            
[124] "province_dhs"                  "province_name"                 "municipality_dhs"             
[127] "municipality_name"             "barangay_name"                 "latitude"                     
[130] "longitude"                     "altitude_gps"                  "altitude_dem"                 
[133] "cc_fips"                       "source"                        "urban_rural"                  
[136] "datum"                         "manila"                        "treated"                      
[139] "birth_year_relative_to_policy" "post_policy"                   "urban"                        
[142] "mother_educ_years"             "number_of_siblings"            "ncr"                          
[145] "num_vaccines"                  "location_group"                "group"                        
  [1] "case_id"                       "cluster_id"                    "household_id"                 
  [4] "sample_weight"                 "mother_age_current"            "residence_type"               
  [7] "household_size"                "household_u5_count"            "region"                       
 [10] "province"                      "municipality"                  "wealth_index"                 
 [13] "wealth_score"                  "birth_index"                   "birth_order"                  
 [16] "is_twin"                       "birth_month"                   "birth_year"                   
 [19] "birth_date"                    "child_sex"                     "child_alive"                  
 [22] "age_at_death"                  "child_age"                     "preceding_birth_interval"     
 [25] "succeeding_birth_interval"     "delivery_place"                "birth_size"                   
 [28] "birth_weight"                  "received_bcg"                  "bcg_day"                      
 [31] "bcg_month"                     "bcg_year"                      "received_dpt1"                
 [34] "dpt1_day"                      "dpt1_month"                    "dpt1_year"                    
 [37] "received_polio1"               "polio1_day"                    "polio1_month"                 
 [40] "polio1_year"                   "received_dpt2"                 "dpt2_day"                     
 [43] "dpt2_month"                    "dpt2_year"                     "received_polio2"              
 [46] "polio2_day"                    "polio2_month"                  "polio2_year"                  
 [49] "received_dpt3"                 "dpt3_day"                      "dpt3_month"                   
 [52] "dpt3_year"                     "received_polio3"               "polio3_day"                   
 [55] "polio3_month"                  "polio3_year"                   "received_measles"             
 [58] "measles_day"                   "measles_month"                 "measles_year"                 
 [61] "ever_vaccinated"               "diarrhea_recent"               "fever_recent"                 
 [64] "cough_recent"                  "total_children_ever"           "sons_at_home"                 
 [67] "daughters_at_home"             "sons_died"                     "daughters_died"               
 [70] "births_last5"                  "births_last1"                  "births_this_month"            
 [73] "mother_age_first_birth"        "currently_pregnant"            "living_children"              
 [76] "living_plus_preg"              "living_plus_preg6"             "contraceptive_method"         
 [79] "contraceptive_intent"          "wanted_last_child"             "mother_educ_level"            
 [82] "mother_literate"               "mother_educ_attain"            "marital_status"               
 [85] "mother_age_first_sex"          "ideal_children"                "partner_educ_level"           
 [88] "partner_occupation"            "mother_occupation"             "mother_work_location"         
 [91] "survey_year"                   "region_dhs.x"                  "admin_region_name.x"          
 [94] "province_dhs.x"                "province_name.x"               "municipality_dhs.x"           
 [97] "municipality_name.x"           "barangay_name.x"               "latitude.x"                   
[100] "longitude.x"                   "altitude_gps.x"                "altitude_dem.x"               
[103] "cc_fips.x"                     "source.x"                      "urban_rural.x"                
[106] "datum.x"                       "region_dhs.y"                  "admin_region_name.y"          
[109] "province_dhs.y"                "province_name.y"               "municipality_dhs.y"           
[112] "municipality_name.y"           "barangay_name.y"               "latitude.y"                   
[115] "longitude.y"                   "altitude_gps.y"                "altitude_dem.y"               
[118] "cc_fips.y"                     "source.y"                      "urban_rural.y"                
[121] "datum.y"                       "region_dhs"                    "admin_region_name"            
[124] "province_dhs"                  "province_name"                 "municipality_dhs"             
[127] "municipality_name"             "barangay_name"                 "latitude"                     
[130] "longitude"                     "altitude_gps"                  "altitude_dem"                 
[133] "cc_fips"                       "source"                        "urban_rural"                  
[136] "datum"                         "manila"                        "treated"                      
[139] "birth_year_relative_to_policy" "post_policy"                   "urban"                        
[142] "mother_educ_years"             "number_of_siblings"            "ncr"                          
[145] "num_vaccines"                  "location_group"                "group"                        
             used    (Mb) gc trigger    (Mb)   max used    (Mb)
Ncells    3222493   172.1    5740976   306.7    5740976   306.7
Vcells 2622053306 20004.7 5428700101 41417.7 5975044060 45586.0
[1] 7
[1] 7
[1] 7

|Variable                 |123                  |136                  |151                  |209                  |3410                 |7437                 |
|:------------------------|:--------------------|:--------------------|:--------------------|:--------------------|:--------------------|:--------------------|
|Birth weight (grams)     |5816.11 \\ (3592.43) |5379.28 \\ (3411.51) |3937.93 \\ (2439.54) |4766.51 \\ (3089.10) |5254.08 \\ (3321.64) |5754.47 \\ (3399.87) |
|1 - Very small           |2 (1.6%)             |NA                   |1 (0.7%)             |10 (4.8%)            |116 (3.4%)           |345 (4.7%)           |
|2 - Smaller than average |25 (20.3%)           |34 (25.0%)           |38 (25.2%)           |35 (16.8%)           |584 (17.2%)          |1256 (17.0%)         |
|3 - Average size         |76 (61.8%)           |84 (61.8%)           |84 (55.6%)           |122 (58.7%)          |1990 (58.6%)         |4397 (59.4%)         |
|4 - Larger than average  |18 (14.6%)           |15 (11.0%)           |18 (11.9%)           |28 (13.5%)           |509 (15.0%)          |970 (13.1%)          |
|5 - Very large           |2 (1.6%)             |3 (2.2%)             |10 (6.6%)            |12 (5.8%)            |184 (5.4%)           |407 (5.5%)           |
|9 - Missing              |NA                   |NA                   |NA                   |1 (0.5%)             |NA                   |18 (0.2%)            |
|8 - Don’t know           |NA                   |NA                   |NA                   |NA                   |14 (0.4%)            |13 (0.2%)            |
|1–2 Vaccines             |1 (0.8%)             |3 (2.2%)             |3 (2.0%)             |1 (0.5%)             |95 (2.8%)            |68 (0.9%)            |
|3–4 Vaccines             |3 (2.4%)             |5 (3.7%)             |5 (3.3%)             |4 (1.9%)             |117 (3.4%)           |136 (1.8%)           |
|5+ Vaccines              |31 (25.2%)           |52 (38.2%)           |29 (19.2%)           |38 (18.2%)           |1141 (33.5%)         |1869 (25.1%)         |
|No Vaccines              |88 (71.5%)           |76 (55.9%)           |114 (75.5%)          |166 (79.4%)          |2057 (60.3%)         |5364 (72.1%)         |

|Variable                 |123                  |136                  |151                  |209                  |3410                 |7437                 |
|:------------------------|:--------------------|:--------------------|:--------------------|:--------------------|:--------------------|:--------------------|
|Birth weight (grams)     |5816.11 \\ (3592.43) |5379.28 \\ (3411.51) |3937.93 \\ (2439.54) |4766.51 \\ (3089.10) |5254.08 \\ (3321.64) |5754.47 \\ (3399.87) |
|1 - Very small           |2 (1.6%)             |NA                   |1 (0.7%)             |10 (4.8%)            |116 (3.4%)           |345 (4.7%)           |
|2 - Smaller than average |25 (20.3%)           |34 (25.0%)           |38 (25.2%)           |35 (16.8%)           |584 (17.2%)          |1256 (17.0%)         |
|3 - Average size         |76 (61.8%)           |84 (61.8%)           |84 (55.6%)           |122 (58.7%)          |1990 (58.6%)         |4397 (59.4%)         |
|4 - Larger than average  |18 (14.6%)           |15 (11.0%)           |18 (11.9%)           |28 (13.5%)           |509 (15.0%)          |970 (13.1%)          |
|5 - Very large           |2 (1.6%)             |3 (2.2%)             |10 (6.6%)            |12 (5.8%)            |184 (5.4%)           |407 (5.5%)           |
|9 - Missing              |NA                   |NA                   |NA                   |1 (0.5%)             |NA                   |18 (0.2%)            |
|8 - Don’t know           |NA                   |NA                   |NA                   |NA                   |14 (0.4%)            |13 (0.2%)            |
|1–2 Vaccines             |1 (0.8%)             |3 (2.2%)             |3 (2.0%)             |1 (0.5%)             |95 (2.8%)            |68 (0.9%)            |
|3–4 Vaccines             |3 (2.4%)             |5 (3.7%)             |5 (3.3%)             |4 (1.9%)             |117 (3.4%)           |136 (1.8%)           |
|5+ Vaccines              |31 (25.2%)           |52 (38.2%)           |29 (19.2%)           |38 (18.2%)           |1141 (33.5%)         |1869 (25.1%)         |
|No Vaccines              |88 (71.5%)           |76 (55.9%)           |114 (75.5%)          |166 (79.4%)          |2057 (60.3%)         |5364 (72.1%)         |
           used  (Mb) gc trigger    (Mb)   max used    (Mb)
Ncells  3222541 172.2    5740976   306.7    5740976   306.7
Vcells 24136690 184.2 4342960081 33134.2 5975044060 45586.0
           used  (Mb) gc trigger    (Mb)   max used    (Mb)
Ncells  3220838 172.1    5740976   306.7    5740976   306.7
Vcells 24001138 183.2 3474368065 26507.4 5975044060 45586.0
 [1] "\\begin{table}[H][!h]"                                          
 [2] "\\centering"                                                    
 [3] "\\caption{Average Birth Weight by Location and Policy Period}"  
 [4] "\\centering"                                                    
 [5] "\\begin{tabular}[t]{lc}"                                        
 [6] "\\toprule"                                                      
 [7] "Group & Birth weight (g)\\\\"                                   
 [8] "\\midrule"                                                      
 [9] "Manila, Pre-Policy & \\makecell{4766.51 \\\\ (3089.10)}\\\\"    
[10] "Manila, Post-Policy & \\makecell{3864.21 \\\\ (2419.37)}\\\\"   
[11] "Other NCR, Pre-Policy & \\makecell{5816.11 \\\\ (3592.43)}\\\\" 
[12] "Other NCR, Post-Policy & \\makecell{5379.28 \\\\ (3411.51)}\\\\"
[13] "Non-NCR, Pre-Policy & \\makecell{5754.47 \\\\ (3399.87)}\\\\"   
[14] "Non-NCR, Post-Policy & \\makecell{5254.08 \\\\ (3321.64)}\\\\"  
[15] "\\bottomrule"                                                   
[16] "\\end{tabular}"                                                 
[17] "\\end{table}"                                                   
 [1] "\\begin{table}[H][!h]"                                          
 [2] "\\centering"                                                    
 [3] "\\caption{Average Birth Weight by Location and Policy Period}"  
 [4] "\\centering"                                                    
 [5] "\\begin{tabular}[t]{lc}"                                        
 [6] "\\toprule"                                                      
 [7] "Group & Birth weight (g)\\\\"                                   
 [8] "\\midrule"                                                      
 [9] "Manila, Pre-Policy & \\makecell{4766.51 \\\\ (3089.10)}\\\\"    
[10] "Manila, Post-Policy & \\makecell{3864.21 \\\\ (2419.37)}\\\\"   
[11] "Other NCR, Pre-Policy & \\makecell{5816.11 \\\\ (3592.43)}\\\\" 
[12] "Other NCR, Post-Policy & \\makecell{5379.28 \\\\ (3411.51)}\\\\"
[13] "Non-NCR, Pre-Policy & \\makecell{5754.47 \\\\ (3399.87)}\\\\"   
[14] "Non-NCR, Post-Policy & \\makecell{5254.08 \\\\ (3321.64)}\\\\"  
[15] "\\bottomrule"                                                   
[16] "\\end{tabular}"                                                 
[17] "\\end{table}"                                                   
           used  (Mb) gc trigger    (Mb)   max used    (Mb)
Ncells  3220342 172.0    5740976   306.7    5740976   306.7
Vcells 24001025 183.2 2779494452 21205.9 5975044060 45586.0
# A tibble: 6 × 136
  case_id           cluster_id household_id sample_weight mother_age_current residence_type household_size
  <chr>                  <dbl>        <dbl>         <dbl>              <dbl> <dbl+lbl>               <dbl>
1 "       1   2  2"          1            2       1452369                 27 1 [urban]                   3
2 "       1   3  2"          1            3       1452369                 23 1 [urban]                   4
3 "       1   3  2"          1            3       1452369                 23 1 [urban]                   4
4 "       1   4  2"          1            4       1452369                 21 1 [urban]                  11
5 "       1   4  2"          1            4       1452369                 21 1 [urban]                  11
6 "       1   4  9"          1            4       1452369                 24 1 [urban]                  11
# ℹ 129 more variables: household_u5_count <dbl>, region <dbl+lbl>, province <dbl+lbl>, municipality <dbl>,
#   wealth_index <dbl+lbl>, wealth_score <dbl>, birth_index <dbl>, birth_order <dbl>, is_twin <dbl+lbl>,
#   birth_month <dbl>, birth_year <dbl>, birth_date <dbl>, child_sex <dbl+lbl>, child_alive <dbl+lbl>,
#   age_at_death <dbl+lbl>, child_age <dbl>, preceding_birth_interval <dbl>, succeeding_birth_interval <dbl>,
#   delivery_place <dbl+lbl>, birth_size <dbl+lbl>, birth_weight <dbl+lbl>, received_bcg <dbl+lbl>,
#   bcg_day <dbl+lbl>, bcg_month <dbl+lbl>, bcg_year <dbl+lbl>, received_dpt1 <dbl+lbl>, dpt1_day <dbl+lbl>,
#   dpt1_month <dbl+lbl>, dpt1_year <dbl+lbl>, received_polio1 <dbl+lbl>, polio1_day <dbl+lbl>, …
# ℹ Use `colnames()` to see all variable names
  [1] "case_id"                   "cluster_id"                "household_id"             
  [4] "sample_weight"             "mother_age_current"        "residence_type"           
  [7] "household_size"            "household_u5_count"        "region"                   
 [10] "province"                  "municipality"              "wealth_index"             
 [13] "wealth_score"              "birth_index"               "birth_order"              
 [16] "is_twin"                   "birth_month"               "birth_year"               
 [19] "birth_date"                "child_sex"                 "child_alive"              
 [22] "age_at_death"              "child_age"                 "preceding_birth_interval" 
 [25] "succeeding_birth_interval" "delivery_place"            "birth_size"               
 [28] "birth_weight"              "received_bcg"              "bcg_day"                  
 [31] "bcg_month"                 "bcg_year"                  "received_dpt1"            
 [34] "dpt1_day"                  "dpt1_month"                "dpt1_year"                
 [37] "received_polio1"           "polio1_day"                "polio1_month"             
 [40] "polio1_year"               "received_dpt2"             "dpt2_day"                 
 [43] "dpt2_month"                "dpt2_year"                 "received_polio2"          
 [46] "polio2_day"                "polio2_month"              "polio2_year"              
 [49] "received_dpt3"             "dpt3_day"                  "dpt3_month"               
 [52] "dpt3_year"                 "received_polio3"           "polio3_day"               
 [55] "polio3_month"              "polio3_year"               "received_measles"         
 [58] "measles_day"               "measles_month"             "measles_year"             
 [61] "ever_vaccinated"           "diarrhea_recent"           "fever_recent"             
 [64] "cough_recent"              "total_children_ever"       "sons_at_home"             
 [67] "daughters_at_home"         "sons_died"                 "daughters_died"           
 [70] "births_last5"              "births_last1"              "births_this_month"        
 [73] "mother_age_first_birth"    "currently_pregnant"        "living_children"          
 [76] "living_plus_preg"          "living_plus_preg6"         "contraceptive_method"     
 [79] "contraceptive_intent"      "wanted_last_child"         "mother_educ_level"        
 [82] "mother_literate"           "mother_educ_attain"        "marital_status"           
 [85] "mother_age_first_sex"      "ideal_children"            "partner_educ_level"       
 [88] "partner_occupation"        "mother_occupation"         "mother_work_location"     
 [91] "survey_year"               "region_dhs.x"              "admin_region_name.x"      
 [94] "province_dhs.x"            "province_name.x"           "municipality_dhs.x"       
 [97] "municipality_name.x"       "barangay_name.x"           "latitude.x"               
[100] "longitude.x"               "altitude_gps.x"            "altitude_dem.x"           
[103] "cc_fips.x"                 "source.x"                  "urban_rural.x"            
[106] "datum.x"                   "region_dhs.y"              "admin_region_name.y"      
[109] "province_dhs.y"            "province_name.y"           "municipality_dhs.y"       
[112] "municipality_name.y"       "barangay_name.y"           "latitude.y"               
[115] "longitude.y"               "altitude_gps.y"            "altitude_dem.y"           
[118] "cc_fips.y"                 "source.y"                  "urban_rural.y"            
[121] "datum.y"                   "region_dhs"                "admin_region_name"        
[124] "province_dhs"              "province_name"             "municipality_dhs"         
[127] "municipality_name"         "barangay_name"             "latitude"                 
[130] "longitude"                 "altitude_gps"              "altitude_dem"             
[133] "cc_fips"                   "source"                    "urban_rural"              
[136] "datum"                    
  [1] "case_id"                   "cluster_id"                "household_id"             
  [4] "sample_weight"             "mother_age_current"        "residence_type"           
  [7] "household_size"            "household_u5_count"        "region"                   
 [10] "province"                  "municipality"              "wealth_index"             
 [13] "wealth_score"              "birth_index"               "birth_order"              
 [16] "is_twin"                   "birth_month"               "birth_year"               
 [19] "birth_date"                "child_sex"                 "child_alive"              
 [22] "age_at_death"              "child_age"                 "preceding_birth_interval" 
 [25] "succeeding_birth_interval" "delivery_place"            "birth_size"               
 [28] "birth_weight"              "received_bcg"              "bcg_day"                  
 [31] "bcg_month"                 "bcg_year"                  "received_dpt1"            
 [34] "dpt1_day"                  "dpt1_month"                "dpt1_year"                
 [37] "received_polio1"           "polio1_day"                "polio1_month"             
 [40] "polio1_year"               "received_dpt2"             "dpt2_day"                 
 [43] "dpt2_month"                "dpt2_year"                 "received_polio2"          
 [46] "polio2_day"                "polio2_month"              "polio2_year"              
 [49] "received_dpt3"             "dpt3_day"                  "dpt3_month"               
 [52] "dpt3_year"                 "received_polio3"           "polio3_day"               
 [55] "polio3_month"              "polio3_year"               "received_measles"         
 [58] "measles_day"               "measles_month"             "measles_year"             
 [61] "ever_vaccinated"           "diarrhea_recent"           "fever_recent"             
 [64] "cough_recent"              "total_children_ever"       "sons_at_home"             
 [67] "daughters_at_home"         "sons_died"                 "daughters_died"           
 [70] "births_last5"              "births_last1"              "births_this_month"        
 [73] "mother_age_first_birth"    "currently_pregnant"        "living_children"          
 [76] "living_plus_preg"          "living_plus_preg6"         "contraceptive_method"     
 [79] "contraceptive_intent"      "wanted_last_child"         "mother_educ_level"        
 [82] "mother_literate"           "mother_educ_attain"        "marital_status"           
 [85] "mother_age_first_sex"      "ideal_children"            "partner_educ_level"       
 [88] "partner_occupation"        "mother_occupation"         "mother_work_location"     
 [91] "survey_year"               "region_dhs.x"              "admin_region_name.x"      
 [94] "province_dhs.x"            "province_name.x"           "municipality_dhs.x"       
 [97] "municipality_name.x"       "barangay_name.x"           "latitude.x"               
[100] "longitude.x"               "altitude_gps.x"            "altitude_dem.x"           
[103] "cc_fips.x"                 "source.x"                  "urban_rural.x"            
[106] "datum.x"                   "region_dhs.y"              "admin_region_name.y"      
[109] "province_dhs.y"            "province_name.y"           "municipality_dhs.y"       
[112] "municipality_name.y"       "barangay_name.y"           "latitude.y"               
[115] "longitude.y"               "altitude_gps.y"            "altitude_dem.y"           
[118] "cc_fips.y"                 "source.y"                  "urban_rural.y"            
[121] "datum.y"                   "region_dhs"                "admin_region_name"        
[124] "province_dhs"              "province_name"             "municipality_dhs"         
[127] "municipality_name"         "barangay_name"             "latitude"                 
[130] "longitude"                 "altitude_gps"              "altitude_dem"             
[133] "cc_fips"                   "source"                    "urban_rural"              
[136] "datum"                    
  [1] "case_id"                   "cluster_id"                "household_id"             
  [4] "sample_weight"             "mother_age_current"        "residence_type"           
  [7] "household_size"            "household_u5_count"        "region"                   
 [10] "province"                  "municipality"              "wealth_index"             
 [13] "wealth_score"              "birth_index"               "birth_order"              
 [16] "is_twin"                   "birth_month"               "birth_year"               
 [19] "birth_date"                "child_sex"                 "child_alive"              
 [22] "age_at_death"              "child_age"                 "preceding_birth_interval" 
 [25] "succeeding_birth_interval" "delivery_place"            "birth_size"               
 [28] "birth_weight"              "received_bcg"              "bcg_day"                  
 [31] "bcg_month"                 "bcg_year"                  "received_dpt1"            
 [34] "dpt1_day"                  "dpt1_month"                "dpt1_year"                
 [37] "received_polio1"           "polio1_day"                "polio1_month"             
 [40] "polio1_year"               "received_dpt2"             "dpt2_day"                 
 [43] "dpt2_month"                "dpt2_year"                 "received_polio2"          
 [46] "polio2_day"                "polio2_month"              "polio2_year"              
 [49] "received_dpt3"             "dpt3_day"                  "dpt3_month"               
 [52] "dpt3_year"                 "received_polio3"           "polio3_day"               
 [55] "polio3_month"              "polio3_year"               "received_measles"         
 [58] "measles_day"               "measles_month"             "measles_year"             
 [61] "ever_vaccinated"           "diarrhea_recent"           "fever_recent"             
 [64] "cough_recent"              "total_children_ever"       "sons_at_home"             
 [67] "daughters_at_home"         "sons_died"                 "daughters_died"           
 [70] "births_last5"              "births_last1"              "births_this_month"        
 [73] "mother_age_first_birth"    "currently_pregnant"        "living_children"          
 [76] "living_plus_preg"          "living_plus_preg6"         "contraceptive_method"     
 [79] "contraceptive_intent"      "wanted_last_child"         "mother_educ_level"        
 [82] "mother_literate"           "mother_educ_attain"        "marital_status"           
 [85] "mother_age_first_sex"      "ideal_children"            "partner_educ_level"       
 [88] "partner_occupation"        "mother_occupation"         "mother_work_location"     
 [91] "survey_year"               "region_dhs.x"              "admin_region_name.x"      
 [94] "province_dhs.x"            "province_name.x"           "municipality_dhs.x"       
 [97] "municipality_name.x"       "barangay_name.x"           "latitude.x"               
[100] "longitude.x"               "altitude_gps.x"            "altitude_dem.x"           
[103] "cc_fips.x"                 "source.x"                  "urban_rural.x"            
[106] "datum.x"                   "region_dhs.y"              "admin_region_name.y"      
[109] "province_dhs.y"            "province_name.y"           "municipality_dhs.y"       
[112] "municipality_name.y"       "barangay_name.y"           "latitude.y"               
[115] "longitude.y"               "altitude_gps.y"            "altitude_dem.y"           
[118] "cc_fips.y"                 "source.y"                  "urban_rural.y"            
[121] "datum.y"                   "region_dhs"                "admin_region_name"        
[124] "province_dhs"              "province_name"             "municipality_dhs"         
[127] "municipality_name"         "barangay_name"             "latitude"                 
[130] "longitude"                 "altitude_gps"              "altitude_dem"             
[133] "cc_fips"                   "source"                    "urban_rural"              
[136] "datum"                    
  [1] "case_id"                   "cluster_id"                "household_id"             
  [4] "sample_weight"             "mother_age_current"        "residence_type"           
  [7] "household_size"            "household_u5_count"        "region"                   
 [10] "province"                  "municipality"              "wealth_index"             
 [13] "wealth_score"              "birth_index"               "birth_order"              
 [16] "is_twin"                   "birth_month"               "birth_year"               
 [19] "birth_date"                "child_sex"                 "child_alive"              
 [22] "age_at_death"              "child_age"                 "preceding_birth_interval" 
 [25] "succeeding_birth_interval" "delivery_place"            "birth_size"               
 [28] "birth_weight"              "received_bcg"              "bcg_day"                  
 [31] "bcg_month"                 "bcg_year"                  "received_dpt1"            
 [34] "dpt1_day"                  "dpt1_month"                "dpt1_year"                
 [37] "received_polio1"           "polio1_day"                "polio1_month"             
 [40] "polio1_year"               "received_dpt2"             "dpt2_day"                 
 [43] "dpt2_month"                "dpt2_year"                 "received_polio2"          
 [46] "polio2_day"                "polio2_month"              "polio2_year"              
 [49] "received_dpt3"             "dpt3_day"                  "dpt3_month"               
 [52] "dpt3_year"                 "received_polio3"           "polio3_day"               
 [55] "polio3_month"              "polio3_year"               "received_measles"         
 [58] "measles_day"               "measles_month"             "measles_year"             
 [61] "ever_vaccinated"           "diarrhea_recent"           "fever_recent"             
 [64] "cough_recent"              "total_children_ever"       "sons_at_home"             
 [67] "daughters_at_home"         "sons_died"                 "daughters_died"           
 [70] "births_last5"              "births_last1"              "births_this_month"        
 [73] "mother_age_first_birth"    "currently_pregnant"        "living_children"          
 [76] "living_plus_preg"          "living_plus_preg6"         "contraceptive_method"     
 [79] "contraceptive_intent"      "wanted_last_child"         "mother_educ_level"        
 [82] "mother_literate"           "mother_educ_attain"        "marital_status"           
 [85] "mother_age_first_sex"      "ideal_children"            "partner_educ_level"       
 [88] "partner_occupation"        "mother_occupation"         "mother_work_location"     
 [91] "survey_year"               "region_dhs.x"              "admin_region_name.x"      
 [94] "province_dhs.x"            "province_name.x"           "municipality_dhs.x"       
 [97] "municipality_name.x"       "barangay_name.x"           "latitude.x"               
[100] "longitude.x"               "altitude_gps.x"            "altitude_dem.x"           
[103] "cc_fips.x"                 "source.x"                  "urban_rural.x"            
[106] "datum.x"                   "region_dhs.y"              "admin_region_name.y"      
[109] "province_dhs.y"            "province_name.y"           "municipality_dhs.y"       
[112] "municipality_name.y"       "barangay_name.y"           "latitude.y"               
[115] "longitude.y"               "altitude_gps.y"            "altitude_dem.y"           
[118] "cc_fips.y"                 "source.y"                  "urban_rural.y"            
[121] "datum.y"                   "region_dhs"                "admin_region_name"        
[124] "province_dhs"              "province_name"             "municipality_dhs"         
[127] "municipality_name"         "barangay_name"             "latitude"                 
[130] "longitude"                 "altitude_gps"              "altitude_dem"             
[133] "cc_fips"                   "source"                    "urban_rural"              
[136] "datum"                    
           used  (Mb) gc trigger    (Mb)   max used    (Mb)
Ncells  3366595 179.8    5740976   306.7    5740976   306.7
Vcells 24254013 185.1 2223595562 16964.7 5975044060 45586.0
  [1] "case_id"                   "cluster_id"                "household_id"             
  [4] "sample_weight"             "mother_age_current"        "residence_type"           
  [7] "household_size"            "household_u5_count"        "region"                   
 [10] "province"                  "municipality"              "wealth_index"             
 [13] "wealth_score"              "birth_index"               "birth_order"              
 [16] "is_twin"                   "birth_month"               "birth_year"               
 [19] "birth_date"                "child_sex"                 "child_alive"              
 [22] "age_at_death"              "child_age"                 "preceding_birth_interval" 
 [25] "succeeding_birth_interval" "delivery_place"            "birth_size"               
 [28] "birth_weight"              "received_bcg"              "bcg_day"                  
 [31] "bcg_month"                 "bcg_year"                  "received_dpt1"            
 [34] "dpt1_day"                  "dpt1_month"                "dpt1_year"                
 [37] "received_polio1"           "polio1_day"                "polio1_month"             
 [40] "polio1_year"               "received_dpt2"             "dpt2_day"                 
 [43] "dpt2_month"                "dpt2_year"                 "received_polio2"          
 [46] "polio2_day"                "polio2_month"              "polio2_year"              
 [49] "received_dpt3"             "dpt3_day"                  "dpt3_month"               
 [52] "dpt3_year"                 "received_polio3"           "polio3_day"               
 [55] "polio3_month"              "polio3_year"               "received_measles"         
 [58] "measles_day"               "measles_month"             "measles_year"             
 [61] "ever_vaccinated"           "diarrhea_recent"           "fever_recent"             
 [64] "cough_recent"              "total_children_ever"       "sons_at_home"             
 [67] "daughters_at_home"         "sons_died"                 "daughters_died"           
 [70] "births_last5"              "births_last1"              "births_this_month"        
 [73] "mother_age_first_birth"    "currently_pregnant"        "living_children"          
 [76] "living_plus_preg"          "living_plus_preg6"         "contraceptive_method"     
 [79] "contraceptive_intent"      "wanted_last_child"         "mother_educ_level"        
 [82] "mother_literate"           "mother_educ_attain"        "marital_status"           
 [85] "mother_age_first_sex"      "ideal_children"            "partner_educ_level"       
 [88] "partner_occupation"        "mother_occupation"         "mother_work_location"     
 [91] "survey_year"               "region_dhs.x"              "admin_region_name.x"      
 [94] "province_dhs.x"            "province_name.x"           "municipality_dhs.x"       
 [97] "municipality_name.x"       "barangay_name.x"           "latitude.x"               
[100] "longitude.x"               "altitude_gps.x"            "altitude_dem.x"           
[103] "cc_fips.x"                 "source.x"                  "urban_rural.x"            
[106] "datum.x"                   "region_dhs.y"              "admin_region_name.y"      
[109] "province_dhs.y"            "province_name.y"           "municipality_dhs.y"       
[112] "municipality_name.y"       "barangay_name.y"           "latitude.y"               
[115] "longitude.y"               "altitude_gps.y"            "altitude_dem.y"           
[118] "cc_fips.y"                 "source.y"                  "urban_rural.y"            
[121] "datum.y"                   "region_dhs"                "admin_region_name"        
[124] "province_dhs"              "province_name"             "municipality_dhs"         
[127] "municipality_name"         "barangay_name"             "latitude"                 
[130] "longitude"                 "altitude_gps"              "altitude_dem"             
[133] "cc_fips"                   "source"                    "urban_rural"              
[136] "datum"                    
  [1] "case_id"                   "cluster_id"                "household_id"             
  [4] "sample_weight"             "mother_age_current"        "residence_type"           
  [7] "household_size"            "household_u5_count"        "region"                   
 [10] "province"                  "municipality"              "wealth_index"             
 [13] "wealth_score"              "birth_index"               "birth_order"              
 [16] "is_twin"                   "birth_month"               "birth_year"               
 [19] "birth_date"                "child_sex"                 "child_alive"              
 [22] "age_at_death"              "child_age"                 "preceding_birth_interval" 
 [25] "succeeding_birth_interval" "delivery_place"            "birth_size"               
 [28] "birth_weight"              "received_bcg"              "bcg_day"                  
 [31] "bcg_month"                 "bcg_year"                  "received_dpt1"            
 [34] "dpt1_day"                  "dpt1_month"                "dpt1_year"                
 [37] "received_polio1"           "polio1_day"                "polio1_month"             
 [40] "polio1_year"               "received_dpt2"             "dpt2_day"                 
 [43] "dpt2_month"                "dpt2_year"                 "received_polio2"          
 [46] "polio2_day"                "polio2_month"              "polio2_year"              
 [49] "received_dpt3"             "dpt3_day"                  "dpt3_month"               
 [52] "dpt3_year"                 "received_polio3"           "polio3_day"               
 [55] "polio3_month"              "polio3_year"               "received_measles"         
 [58] "measles_day"               "measles_month"             "measles_year"             
 [61] "ever_vaccinated"           "diarrhea_recent"           "fever_recent"             
 [64] "cough_recent"              "total_children_ever"       "sons_at_home"             
 [67] "daughters_at_home"         "sons_died"                 "daughters_died"           
 [70] "births_last5"              "births_last1"              "births_this_month"        
 [73] "mother_age_first_birth"    "currently_pregnant"        "living_children"          
 [76] "living_plus_preg"          "living_plus_preg6"         "contraceptive_method"     
 [79] "contraceptive_intent"      "wanted_last_child"         "mother_educ_level"        
 [82] "mother_literate"           "mother_educ_attain"        "marital_status"           
 [85] "mother_age_first_sex"      "ideal_children"            "partner_educ_level"       
 [88] "partner_occupation"        "mother_occupation"         "mother_work_location"     
 [91] "survey_year"               "region_dhs.x"              "admin_region_name.x"      
 [94] "province_dhs.x"            "province_name.x"           "municipality_dhs.x"       
 [97] "municipality_name.x"       "barangay_name.x"           "latitude.x"               
[100] "longitude.x"               "altitude_gps.x"            "altitude_dem.x"           
[103] "cc_fips.x"                 "source.x"                  "urban_rural.x"            
[106] "datum.x"                   "region_dhs.y"              "admin_region_name.y"      
[109] "province_dhs.y"            "province_name.y"           "municipality_dhs.y"       
[112] "municipality_name.y"       "barangay_name.y"           "latitude.y"               
[115] "longitude.y"               "altitude_gps.y"            "altitude_dem.y"           
[118] "cc_fips.y"                 "source.y"                  "urban_rural.y"            
[121] "datum.y"                   "region_dhs"                "admin_region_name"        
[124] "province_dhs"              "province_name"             "municipality_dhs"         
[127] "municipality_name"         "barangay_name"             "latitude"                 
[130] "longitude"                 "altitude_gps"              "altitude_dem"             
[133] "cc_fips"                   "source"                    "urban_rural"              
[136] "datum"                    
  [1] "case_id"                   "cluster_id"                "household_id"             
  [4] "sample_weight"             "mother_age_current"        "residence_type"           
  [7] "household_size"            "household_u5_count"        "region"                   
 [10] "province"                  "municipality"              "wealth_index"             
 [13] "wealth_score"              "birth_index"               "birth_order"              
 [16] "is_twin"                   "birth_month"               "birth_year"               
 [19] "birth_date"                "child_sex"                 "child_alive"              
 [22] "age_at_death"              "child_age"                 "preceding_birth_interval" 
 [25] "succeeding_birth_interval" "delivery_place"            "birth_size"               
 [28] "birth_weight"              "received_bcg"              "bcg_day"                  
 [31] "bcg_month"                 "bcg_year"                  "received_dpt1"            
 [34] "dpt1_day"                  "dpt1_month"                "dpt1_year"                
 [37] "received_polio1"           "polio1_day"                "polio1_month"             
 [40] "polio1_year"               "received_dpt2"             "dpt2_day"                 
 [43] "dpt2_month"                "dpt2_year"                 "received_polio2"          
 [46] "polio2_day"                "polio2_month"              "polio2_year"              
 [49] "received_dpt3"             "dpt3_day"                  "dpt3_month"               
 [52] "dpt3_year"                 "received_polio3"           "polio3_day"               
 [55] "polio3_month"              "polio3_year"               "received_measles"         
 [58] "measles_day"               "measles_month"             "measles_year"             
 [61] "ever_vaccinated"           "diarrhea_recent"           "fever_recent"             
 [64] "cough_recent"              "total_children_ever"       "sons_at_home"             
 [67] "daughters_at_home"         "sons_died"                 "daughters_died"           
 [70] "births_last5"              "births_last1"              "births_this_month"        
 [73] "mother_age_first_birth"    "currently_pregnant"        "living_children"          
 [76] "living_plus_preg"          "living_plus_preg6"         "contraceptive_method"     
 [79] "contraceptive_intent"      "wanted_last_child"         "mother_educ_level"        
 [82] "mother_literate"           "mother_educ_attain"        "marital_status"           
 [85] "mother_age_first_sex"      "ideal_children"            "partner_educ_level"       
 [88] "partner_occupation"        "mother_occupation"         "mother_work_location"     
 [91] "survey_year"               "region_dhs.x"              "admin_region_name.x"      
 [94] "province_dhs.x"            "province_name.x"           "municipality_dhs.x"       
 [97] "municipality_name.x"       "barangay_name.x"           "latitude.x"               
[100] "longitude.x"               "altitude_gps.x"            "altitude_dem.x"           
[103] "cc_fips.x"                 "source.x"                  "urban_rural.x"            
[106] "datum.x"                   "region_dhs.y"              "admin_region_name.y"      
[109] "province_dhs.y"            "province_name.y"           "municipality_dhs.y"       
[112] "municipality_name.y"       "barangay_name.y"           "latitude.y"               
[115] "longitude.y"               "altitude_gps.y"            "altitude_dem.y"           
[118] "cc_fips.y"                 "source.y"                  "urban_rural.y"            
[121] "datum.y"                   "region_dhs"                "admin_region_name"        
[124] "province_dhs"              "province_name"             "municipality_dhs"         
[127] "municipality_name"         "barangay_name"             "latitude"                 
[130] "longitude"                 "altitude_gps"              "altitude_dem"             
[133] "cc_fips"                   "source"                    "urban_rural"              
[136] "datum"                    
  [1] "case_id"                   "cluster_id"                "household_id"             
  [4] "sample_weight"             "mother_age_current"        "residence_type"           
  [7] "household_size"            "household_u5_count"        "region"                   
 [10] "province"                  "municipality"              "wealth_index"             
 [13] "wealth_score"              "birth_index"               "birth_order"              
 [16] "is_twin"                   "birth_month"               "birth_year"               
 [19] "birth_date"                "child_sex"                 "child_alive"              
 [22] "age_at_death"              "child_age"                 "preceding_birth_interval" 
 [25] "succeeding_birth_interval" "delivery_place"            "birth_size"               
 [28] "birth_weight"              "received_bcg"              "bcg_day"                  
 [31] "bcg_month"                 "bcg_year"                  "received_dpt1"            
 [34] "dpt1_day"                  "dpt1_month"                "dpt1_year"                
 [37] "received_polio1"           "polio1_day"                "polio1_month"             
 [40] "polio1_year"               "received_dpt2"             "dpt2_day"                 
 [43] "dpt2_month"                "dpt2_year"                 "received_polio2"          
 [46] "polio2_day"                "polio2_month"              "polio2_year"              
 [49] "received_dpt3"             "dpt3_day"                  "dpt3_month"               
 [52] "dpt3_year"                 "received_polio3"           "polio3_day"               
 [55] "polio3_month"              "polio3_year"               "received_measles"         
 [58] "measles_day"               "measles_month"             "measles_year"             
 [61] "ever_vaccinated"           "diarrhea_recent"           "fever_recent"             
 [64] "cough_recent"              "total_children_ever"       "sons_at_home"             
 [67] "daughters_at_home"         "sons_died"                 "daughters_died"           
 [70] "births_last5"              "births_last1"              "births_this_month"        
 [73] "mother_age_first_birth"    "currently_pregnant"        "living_children"          
 [76] "living_plus_preg"          "living_plus_preg6"         "contraceptive_method"     
 [79] "contraceptive_intent"      "wanted_last_child"         "mother_educ_level"        
 [82] "mother_literate"           "mother_educ_attain"        "marital_status"           
 [85] "mother_age_first_sex"      "ideal_children"            "partner_educ_level"       
 [88] "partner_occupation"        "mother_occupation"         "mother_work_location"     
 [91] "survey_year"               "region_dhs.x"              "admin_region_name.x"      
 [94] "province_dhs.x"            "province_name.x"           "municipality_dhs.x"       
 [97] "municipality_name.x"       "barangay_name.x"           "latitude.x"               
[100] "longitude.x"               "altitude_gps.x"            "altitude_dem.x"           
[103] "cc_fips.x"                 "source.x"                  "urban_rural.x"            
[106] "datum.x"                   "region_dhs.y"              "admin_region_name.y"      
[109] "province_dhs.y"            "province_name.y"           "municipality_dhs.y"       
[112] "municipality_name.y"       "barangay_name.y"           "latitude.y"               
[115] "longitude.y"               "altitude_gps.y"            "altitude_dem.y"           
[118] "cc_fips.y"                 "source.y"                  "urban_rural.y"            
[121] "datum.y"                   "region_dhs"                "admin_region_name"        
[124] "province_dhs"              "province_name"             "municipality_dhs"         
[127] "municipality_name"         "barangay_name"             "latitude"                 
[130] "longitude"                 "altitude_gps"              "altitude_dem"             
[133] "cc_fips"                   "source"                    "urban_rural"              
[136] "datum"                    
  [1] "case_id"                   "cluster_id"                "household_id"             
  [4] "sample_weight"             "mother_age_current"        "residence_type"           
  [7] "household_size"            "household_u5_count"        "region"                   
 [10] "province"                  "municipality"              "wealth_index"             
 [13] "wealth_score"              "birth_index"               "birth_order"              
 [16] "is_twin"                   "birth_month"               "birth_year"               
 [19] "birth_date"                "child_sex"                 "child_alive"              
 [22] "age_at_death"              "child_age"                 "preceding_birth_interval" 
 [25] "succeeding_birth_interval" "delivery_place"            "birth_size"               
 [28] "birth_weight"              "received_bcg"              "bcg_day"                  
 [31] "bcg_month"                 "bcg_year"                  "received_dpt1"            
 [34] "dpt1_day"                  "dpt1_month"                "dpt1_year"                
 [37] "received_polio1"           "polio1_day"                "polio1_month"             
 [40] "polio1_year"               "received_dpt2"             "dpt2_day"                 
 [43] "dpt2_month"                "dpt2_year"                 "received_polio2"          
 [46] "polio2_day"                "polio2_month"              "polio2_year"              
 [49] "received_dpt3"             "dpt3_day"                  "dpt3_month"               
 [52] "dpt3_year"                 "received_polio3"           "polio3_day"               
 [55] "polio3_month"              "polio3_year"               "received_measles"         
 [58] "measles_day"               "measles_month"             "measles_year"             
 [61] "ever_vaccinated"           "diarrhea_recent"           "fever_recent"             
 [64] "cough_recent"              "total_children_ever"       "sons_at_home"             
 [67] "daughters_at_home"         "sons_died"                 "daughters_died"           
 [70] "births_last5"              "births_last1"              "births_this_month"        
 [73] "mother_age_first_birth"    "currently_pregnant"        "living_children"          
 [76] "living_plus_preg"          "living_plus_preg6"         "contraceptive_method"     
 [79] "contraceptive_intent"      "wanted_last_child"         "mother_educ_level"        
 [82] "mother_literate"           "mother_educ_attain"        "marital_status"           
 [85] "mother_age_first_sex"      "ideal_children"            "partner_educ_level"       
 [88] "partner_occupation"        "mother_occupation"         "mother_work_location"     
 [91] "survey_year"               "region_dhs.x"              "admin_region_name.x"      
 [94] "province_dhs.x"            "province_name.x"           "municipality_dhs.x"       
 [97] "municipality_name.x"       "barangay_name.x"           "latitude.x"               
[100] "longitude.x"               "altitude_gps.x"            "altitude_dem.x"           
[103] "cc_fips.x"                 "source.x"                  "urban_rural.x"            
[106] "datum.x"                   "region_dhs.y"              "admin_region_name.y"      
[109] "province_dhs.y"            "province_name.y"           "municipality_dhs.y"       
[112] "municipality_name.y"       "barangay_name.y"           "latitude.y"               
[115] "longitude.y"               "altitude_gps.y"            "altitude_dem.y"           
[118] "cc_fips.y"                 "source.y"                  "urban_rural.y"            
[121] "datum.y"                   "region_dhs"                "admin_region_name"        
[124] "province_dhs"              "province_name"             "municipality_dhs"         
[127] "municipality_name"         "barangay_name"             "latitude"                 
[130] "longitude"                 "altitude_gps"              "altitude_dem"             
[133] "cc_fips"                   "source"                    "urban_rural"              
[136] "datum"                    
  [1] "case_id"                   "cluster_id"                "household_id"             
  [4] "sample_weight"             "mother_age_current"        "residence_type"           
  [7] "household_size"            "household_u5_count"        "region"                   
 [10] "province"                  "municipality"              "wealth_index"             
 [13] "wealth_score"              "birth_index"               "birth_order"              
 [16] "is_twin"                   "birth_month"               "birth_year"               
 [19] "birth_date"                "child_sex"                 "child_alive"              
 [22] "age_at_death"              "child_age"                 "preceding_birth_interval" 
 [25] "succeeding_birth_interval" "delivery_place"            "birth_size"               
 [28] "birth_weight"              "received_bcg"              "bcg_day"                  
 [31] "bcg_month"                 "bcg_year"                  "received_dpt1"            
 [34] "dpt1_day"                  "dpt1_month"                "dpt1_year"                
 [37] "received_polio1"           "polio1_day"                "polio1_month"             
 [40] "polio1_year"               "received_dpt2"             "dpt2_day"                 
 [43] "dpt2_month"                "dpt2_year"                 "received_polio2"          
 [46] "polio2_day"                "polio2_month"              "polio2_year"              
 [49] "received_dpt3"             "dpt3_day"                  "dpt3_month"               
 [52] "dpt3_year"                 "received_polio3"           "polio3_day"               
 [55] "polio3_month"              "polio3_year"               "received_measles"         
 [58] "measles_day"               "measles_month"             "measles_year"             
 [61] "ever_vaccinated"           "diarrhea_recent"           "fever_recent"             
 [64] "cough_recent"              "total_children_ever"       "sons_at_home"             
 [67] "daughters_at_home"         "sons_died"                 "daughters_died"           
 [70] "births_last5"              "births_last1"              "births_this_month"        
 [73] "mother_age_first_birth"    "currently_pregnant"        "living_children"          
 [76] "living_plus_preg"          "living_plus_preg6"         "contraceptive_method"     
 [79] "contraceptive_intent"      "wanted_last_child"         "mother_educ_level"        
 [82] "mother_literate"           "mother_educ_attain"        "marital_status"           
 [85] "mother_age_first_sex"      "ideal_children"            "partner_educ_level"       
 [88] "partner_occupation"        "mother_occupation"         "mother_work_location"     
 [91] "survey_year"               "region_dhs.x"              "admin_region_name.x"      
 [94] "province_dhs.x"            "province_name.x"           "municipality_dhs.x"       
 [97] "municipality_name.x"       "barangay_name.x"           "latitude.x"               
[100] "longitude.x"               "altitude_gps.x"            "altitude_dem.x"           
[103] "cc_fips.x"                 "source.x"                  "urban_rural.x"            
[106] "datum.x"                   "region_dhs.y"              "admin_region_name.y"      
[109] "province_dhs.y"            "province_name.y"           "municipality_dhs.y"       
[112] "municipality_name.y"       "barangay_name.y"           "latitude.y"               
[115] "longitude.y"               "altitude_gps.y"            "altitude_dem.y"           
[118] "cc_fips.y"                 "source.y"                  "urban_rural.y"            
[121] "datum.y"                   "region_dhs"                "admin_region_name"        
[124] "province_dhs"              "province_name"             "municipality_dhs"         
[127] "municipality_name"         "barangay_name"             "latitude"                 
[130] "longitude"                 "altitude_gps"              "altitude_dem"             
[133] "cc_fips"                   "source"                    "urban_rural"              
[136] "datum"                    
  [1] "case_id"                   "cluster_id"                "household_id"             
  [4] "sample_weight"             "mother_age_current"        "residence_type"           
  [7] "household_size"            "household_u5_count"        "region"                   
 [10] "province"                  "municipality"              "wealth_index"             
 [13] "wealth_score"              "birth_index"               "birth_order"              
 [16] "is_twin"                   "birth_month"               "birth_year"               
 [19] "birth_date"                "child_sex"                 "child_alive"              
 [22] "age_at_death"              "child_age"                 "preceding_birth_interval" 
 [25] "succeeding_birth_interval" "delivery_place"            "birth_size"               
 [28] "birth_weight"              "received_bcg"              "bcg_day"                  
 [31] "bcg_month"                 "bcg_year"                  "received_dpt1"            
 [34] "dpt1_day"                  "dpt1_month"                "dpt1_year"                
 [37] "received_polio1"           "polio1_day"                "polio1_month"             
 [40] "polio1_year"               "received_dpt2"             "dpt2_day"                 
 [43] "dpt2_month"                "dpt2_year"                 "received_polio2"          
 [46] "polio2_day"                "polio2_month"              "polio2_year"              
 [49] "received_dpt3"             "dpt3_day"                  "dpt3_month"               
 [52] "dpt3_year"                 "received_polio3"           "polio3_day"               
 [55] "polio3_month"              "polio3_year"               "received_measles"         
 [58] "measles_day"               "measles_month"             "measles_year"             
 [61] "ever_vaccinated"           "diarrhea_recent"           "fever_recent"             
 [64] "cough_recent"              "total_children_ever"       "sons_at_home"             
 [67] "daughters_at_home"         "sons_died"                 "daughters_died"           
 [70] "births_last5"              "births_last1"              "births_this_month"        
 [73] "mother_age_first_birth"    "currently_pregnant"        "living_children"          
 [76] "living_plus_preg"          "living_plus_preg6"         "contraceptive_method"     
 [79] "contraceptive_intent"      "wanted_last_child"         "mother_educ_level"        
 [82] "mother_literate"           "mother_educ_attain"        "marital_status"           
 [85] "mother_age_first_sex"      "ideal_children"            "partner_educ_level"       
 [88] "partner_occupation"        "mother_occupation"         "mother_work_location"     
 [91] "survey_year"               "region_dhs.x"              "admin_region_name.x"      
 [94] "province_dhs.x"            "province_name.x"           "municipality_dhs.x"       
 [97] "municipality_name.x"       "barangay_name.x"           "latitude.x"               
[100] "longitude.x"               "altitude_gps.x"            "altitude_dem.x"           
[103] "cc_fips.x"                 "source.x"                  "urban_rural.x"            
[106] "datum.x"                   "region_dhs.y"              "admin_region_name.y"      
[109] "province_dhs.y"            "province_name.y"           "municipality_dhs.y"       
[112] "municipality_name.y"       "barangay_name.y"           "latitude.y"               
[115] "longitude.y"               "altitude_gps.y"            "altitude_dem.y"           
[118] "cc_fips.y"                 "source.y"                  "urban_rural.y"            
[121] "datum.y"                   "region_dhs"                "admin_region_name"        
[124] "province_dhs"              "province_name"             "municipality_dhs"         
[127] "municipality_name"         "barangay_name"             "latitude"                 
[130] "longitude"                 "altitude_gps"              "altitude_dem"             
[133] "cc_fips"                   "source"                    "urban_rural"              
[136] "datum"                    
  [1] "case_id"                   "cluster_id"                "household_id"             
  [4] "sample_weight"             "mother_age_current"        "residence_type"           
  [7] "household_size"            "household_u5_count"        "region"                   
 [10] "province"                  "municipality"              "wealth_index"             
 [13] "wealth_score"              "birth_index"               "birth_order"              
 [16] "is_twin"                   "birth_month"               "birth_year"               
 [19] "birth_date"                "child_sex"                 "child_alive"              
 [22] "age_at_death"              "child_age"                 "preceding_birth_interval" 
 [25] "succeeding_birth_interval" "delivery_place"            "birth_size"               
 [28] "birth_weight"              "received_bcg"              "bcg_day"                  
 [31] "bcg_month"                 "bcg_year"                  "received_dpt1"            
 [34] "dpt1_day"                  "dpt1_month"                "dpt1_year"                
 [37] "received_polio1"           "polio1_day"                "polio1_month"             
 [40] "polio1_year"               "received_dpt2"             "dpt2_day"                 
 [43] "dpt2_month"                "dpt2_year"                 "received_polio2"          
 [46] "polio2_day"                "polio2_month"              "polio2_year"              
 [49] "received_dpt3"             "dpt3_day"                  "dpt3_month"               
 [52] "dpt3_year"                 "received_polio3"           "polio3_day"               
 [55] "polio3_month"              "polio3_year"               "received_measles"         
 [58] "measles_day"               "measles_month"             "measles_year"             
 [61] "ever_vaccinated"           "diarrhea_recent"           "fever_recent"             
 [64] "cough_recent"              "total_children_ever"       "sons_at_home"             
 [67] "daughters_at_home"         "sons_died"                 "daughters_died"           
 [70] "births_last5"              "births_last1"              "births_this_month"        
 [73] "mother_age_first_birth"    "currently_pregnant"        "living_children"          
 [76] "living_plus_preg"          "living_plus_preg6"         "contraceptive_method"     
 [79] "contraceptive_intent"      "wanted_last_child"         "mother_educ_level"        
 [82] "mother_literate"           "mother_educ_attain"        "marital_status"           
 [85] "mother_age_first_sex"      "ideal_children"            "partner_educ_level"       
 [88] "partner_occupation"        "mother_occupation"         "mother_work_location"     
 [91] "survey_year"               "region_dhs.x"              "admin_region_name.x"      
 [94] "province_dhs.x"            "province_name.x"           "municipality_dhs.x"       
 [97] "municipality_name.x"       "barangay_name.x"           "latitude.x"               
[100] "longitude.x"               "altitude_gps.x"            "altitude_dem.x"           
[103] "cc_fips.x"                 "source.x"                  "urban_rural.x"            
[106] "datum.x"                   "region_dhs.y"              "admin_region_name.y"      
[109] "province_dhs.y"            "province_name.y"           "municipality_dhs.y"       
[112] "municipality_name.y"       "barangay_name.y"           "latitude.y"               
[115] "longitude.y"               "altitude_gps.y"            "altitude_dem.y"           
[118] "cc_fips.y"                 "source.y"                  "urban_rural.y"            
[121] "datum.y"                   "region_dhs"                "admin_region_name"        
[124] "province_dhs"              "province_name"             "municipality_dhs"         
[127] "municipality_name"         "barangay_name"             "latitude"                 
[130] "longitude"                 "altitude_gps"              "altitude_dem"             
[133] "cc_fips"                   "source"                    "urban_rural"              
[136] "datum"