### 2025-08-07 data-analysis/outputs/first_stage_sibship.tex

\begin{table}
\centering
\begin{talltblr}[         %% tabularray outer open
entry=none,label=none,
note{}={+ p \num{< 0.1}, * p \num{< 0.05}, ** p \num{< 0.01}, *** p \num{< 0.001}},
note{ }={All regressions include survey year fixed effects.},
]                     %% tabularray outer close
{                     %% tabularray inner open
colspec={Q[]Q[]Q[]},
column{2,3}={}{halign=c,},
column{1}={}{halign=l,},
hline{10}={1,2,3}{solid, black, 0.05em},
}                     %% tabularray inner close
\toprule
& (1) & (2) \\ \midrule %% TinyTableHeader
Manila × Post & \num{0.231} & \num{0.100}+ \\
& (\num{0.227}) & (\num{0.077}) \\
Mother's Education & \num{-0.393}*** & \num{-0.315}*** \\
& (\num{<0.001}) & (\num{<0.001}) \\
Mother's Age & \num{0.230}** & \num{0.207}** \\
& (\num{0.003}) & (\num{0.001}) \\
Wealth Index & \num{-0.287}*** & \num{-0.247}** \\
& (\num{<0.001}) & (\num{0.002}) \\
Num.Obs. & \num{20823} & \num{20823} \\
R2 & \num{0.539} & \num{0.514} \\
R2 Adj. & \num{0.539} & \num{0.513} \\
R2 Within & \num{0.535} & \num{0.511} \\
R2 Within Adj. & \num{0.535} & \num{0.511} \\
RMSE & \num{1.65} & \num{1.53} \\
FE: survey_year & X & X \\
\bottomrule
\end{talltblr}
\end{table} 
\begin{table}
\centering
\begin{talltblr}[         %% tabularray outer open
entry=none,label=none,
note{}={+ p \num{< 0.1}, * p \num{< 0.05}, ** p \num{< 0.01}, *** p \num{< 0.001}},
note{ }={All regressions include survey year fixed effects.},
]                     %% tabularray outer close
{                     %% tabularray inner open
colspec={Q[]Q[]Q[]},
column{2,3}={}{halign=c,},
column{1}={}{halign=l,},
hline{10}={1,2,3}{solid, black, 0.05em},
}                     %% tabularray inner close
\toprule
& (1) & (2) \\ \midrule %% TinyTableHeader
Manila × Post & \num{0.231} & \num{0.100}+ \\
& (\num{0.227}) & (\num{0.077}) \\
Mother's Education & \num{-0.393}*** & \num{-0.315}*** \\
& (\num{<0.001}) & (\num{<0.001}) \\
Mother's Age & \num{0.230}** & \num{0.207}** \\
& (\num{0.003}) & (\num{0.001}) \\
Wealth Index & \num{-0.287}*** & \num{-0.247}** \\
& (\num{<0.001}) & (\num{0.002}) \\
Num.Obs. & \num{20823} & \num{20823} \\
R2 & \num{0.539} & \num{0.514} \\
R2 Adj. & \num{0.539} & \num{0.513} \\
R2 Within & \num{0.535} & \num{0.511} \\
R2 Within Adj. & \num{0.535} & \num{0.511} \\
FE: survey\_year & X & X \\
\bottomrule
\end{talltblr}
\end{table} 
\begin{table}
\centering
\begin{talltblr}[         %% tabularray outer open
entry=none,label=none,
note{}={+ p \num{< 0.1}, * p \num{< 0.05}, ** p \num{< 0.01}, *** p \num{< 0.001}},
note{ }={All regressions include survey year fixed effects.},
]                     %% tabularray outer close
{                     %% tabularray inner open
colspec={Q[]Q[]Q[]},
column{2,3}={}{halign=c,},
column{1}={}{halign=l,},
hline{10}={1,2,3}{solid, black, 0.05em},
}                     %% tabularray inner close
\toprule
& (1) & (2) \\ \midrule %% TinyTableHeader
Manila × Post & \num{0.100}+ & \num{0.231} \\
& (\num{0.077}) & (\num{0.227}) \\
Mother's Education & \num{-0.315}*** & \num{-0.393}*** \\
& (\num{<0.001}) & (\num{<0.001}) \\
Mother's Age & \num{0.207}** & \num{0.230}** \\
& (\num{0.001}) & (\num{0.003}) \\
Wealth Index & \num{-0.247}** & \num{-0.287}*** \\
& (\num{0.002}) & (\num{<0.001}) \\
Num.Obs. & \num{20823} & \num{20823} \\
R2 & \num{0.514} & \num{0.539} \\
R2 Adj. & \num{0.513} & \num{0.539} \\
R2 Within & \num{0.511} & \num{0.535} \\
R2 Within Adj. & \num{0.511} & \num{0.535} \\
FE: survey\_year & X & X \\
\bottomrule
\end{talltblr}
\end{table} 
\begin{table}
\centering
\begin{talltblr}[         %% tabularray outer open
entry=none,label=none,
note{}={+ p \num{< 0.1}, * p \num{< 0.05}, ** p \num{< 0.01}, *** p \num{< 0.001}},
note{ }={All regressions include survey year fixed effects.},
]                     %% tabularray outer close
{                     %% tabularray inner open
colspec={Q[]Q[]Q[]},
column{2,3}={}{halign=c,},
column{1}={}{halign=l,},
hline{10}={1,2,3}{solid, black, 0.05em},
}                     %% tabularray inner close
\toprule
& (1) & (2) \\ \midrule %% TinyTableHeader
Manila × Post & \num{0.100}+ & \num{0.231} \\
& (\num{0.077}) & (\num{0.227}) \\
Mother's Education & \num{-0.315}*** & \num{-0.393}*** \\
& (\num{<0.001}) & (\num{<0.001}) \\
Mother's Age & \num{0.207}** & \num{0.230}** \\
& (\num{0.001}) & (\num{0.003}) \\
Wealth Index & \num{-0.247}** & \num{-0.287}*** \\
& (\num{0.002}) & (\num{<0.001}) \\
Num.Obs. & \num{20823} & \num{20823} \\
R2 & \num{0.514} & \num{0.539} \\
R2 Adj. & \num{0.513} & \num{0.539} \\
R2 Within & \num{0.511} & \num{0.535} \\
R2 Within Adj. & \num{0.511} & \num{0.535} \\
FE: survey\_year & X & X \\
\bottomrule
\end{talltblr}
\end{table} 
          used  (Mb) gc trigger  (Mb) max used  (Mb)
Ncells 2386672 127.5    3882202 207.4  3882202 207.4
Vcells 8326641  63.6   26574801 202.8 26574145 202.8
  [1] "case_id"                   "cluster_id"                "household_id"             
  [4] "sample_weight"             "mother_age_current"        "residence_type"           
  [7] "household_size"            "household_u5_count"        "region"                   
 [10] "province"                  "municipality"              "wealth_index"             
 [13] "wealth_score"              "birth_index"               "birth_order"              
 [16] "is_twin"                   "birth_month"               "birth_year"               
 [19] "birth_date"                "child_sex"                 "child_alive"              
 [22] "age_at_death"              "child_age"                 "preceding_birth_interval" 
 [25] "succeeding_birth_interval" "delivery_place"            "birth_size"               
 [28] "birth_weight"              "received_bcg"              "bcg_day"                  
 [31] "bcg_month"                 "bcg_year"                  "received_dpt1"            
 [34] "dpt1_day"                  "dpt1_month"                "dpt1_year"                
 [37] "received_polio1"           "polio1_day"                "polio1_month"             
 [40] "polio1_year"               "received_dpt2"             "dpt2_day"                 
 [43] "dpt2_month"                "dpt2_year"                 "received_polio2"          
 [46] "polio2_day"                "polio2_month"              "polio2_year"              
 [49] "received_dpt3"             "dpt3_day"                  "dpt3_month"               
 [52] "dpt3_year"                 "received_polio3"           "polio3_day"               
 [55] "polio3_month"              "polio3_year"               "received_measles"         
 [58] "measles_day"               "measles_month"             "measles_year"             
 [61] "ever_vaccinated"           "diarrhea_recent"           "fever_recent"             
 [64] "cough_recent"              "total_children_ever"       "sons_at_home"             
 [67] "daughters_at_home"         "sons_died"                 "daughters_died"           
 [70] "births_last5"              "births_last1"              "births_this_month"        
 [73] "mother_age_first_birth"    "currently_pregnant"        "living_children"          
 [76] "living_plus_preg"          "living_plus_preg6"         "contraceptive_method"     
 [79] "contraceptive_intent"      "wanted_last_child"         "mother_educ_level"        
 [82] "mother_literate"           "mother_educ_attain"        "marital_status"           
 [85] "mother_age_first_sex"      "ideal_children"            "partner_educ_level"       
 [88] "partner_occupation"        "mother_occupation"         "mother_work_location"     
 [91] "survey_year"               "region_dhs.x"              "admin_region_name.x"      
 [94] "province_dhs.x"            "province_name.x"           "municipality_dhs.x"       
 [97] "municipality_name.x"       "barangay_name.x"           "latitude.x"               
[100] "longitude.x"               "altitude_gps.x"            "altitude_dem.x"           
[103] "cc_fips.x"                 "source.x"                  "urban_rural.x"            
[106] "datum.x"                   "region_dhs.y"              "admin_region_name.y"      
[109] "province_dhs.y"            "province_name.y"           "municipality_dhs.y"       
[112] "municipality_name.y"       "barangay_name.y"           "latitude.y"               
[115] "longitude.y"               "altitude_gps.y"            "altitude_dem.y"           
[118] "cc_fips.y"                 "source.y"                  "urban_rural.y"            
[121] "datum.y"                   "region_dhs"                "admin_region_name"        
[124] "province_dhs"              "province_name"             "municipality_dhs"         
[127] "municipality_name"         "barangay_name"             "latitude"                 
[130] "longitude"                 "altitude_gps"              "altitude_dem"             
[133] "cc_fips"                   "source"                    "urban_rural"              
[136] "datum"                     "number_of_siblings"        "manila"                   
[139] "post_policy"               "manila_post"              
  [1] "case_id"                   "cluster_id"                "household_id"             
  [4] "sample_weight"             "mother_age_current"        "residence_type"           
  [7] "household_size"            "household_u5_count"        "region"                   
 [10] "province"                  "municipality"              "wealth_index"             
 [13] "wealth_score"              "birth_index"               "birth_order"              
 [16] "is_twin"                   "birth_month"               "birth_year"               
 [19] "birth_date"                "child_sex"                 "child_alive"              
 [22] "age_at_death"              "child_age"                 "preceding_birth_interval" 
 [25] "succeeding_birth_interval" "delivery_place"            "birth_size"               
 [28] "birth_weight"              "received_bcg"              "bcg_day"                  
 [31] "bcg_month"                 "bcg_year"                  "received_dpt1"            
 [34] "dpt1_day"                  "dpt1_month"                "dpt1_year"                
 [37] "received_polio1"           "polio1_day"                "polio1_month"             
 [40] "polio1_year"               "received_dpt2"             "dpt2_day"                 
 [43] "dpt2_month"                "dpt2_year"                 "received_polio2"          
 [46] "polio2_day"                "polio2_month"              "polio2_year"              
 [49] "received_dpt3"             "dpt3_day"                  "dpt3_month"               
 [52] "dpt3_year"                 "received_polio3"           "polio3_day"               
 [55] "polio3_month"              "polio3_year"               "received_measles"         
 [58] "measles_day"               "measles_month"             "measles_year"             
 [61] "ever_vaccinated"           "diarrhea_recent"           "fever_recent"             
 [64] "cough_recent"              "total_children_ever"       "sons_at_home"             
 [67] "daughters_at_home"         "sons_died"                 "daughters_died"           
 [70] "births_last5"              "births_last1"              "births_this_month"        
 [73] "mother_age_first_birth"    "currently_pregnant"        "living_children"          
 [76] "living_plus_preg"          "living_plus_preg6"         "contraceptive_method"     
 [79] "contraceptive_intent"      "wanted_last_child"         "mother_educ_level"        
 [82] "mother_literate"           "mother_educ_attain"        "marital_status"           
 [85] "mother_age_first_sex"      "ideal_children"            "partner_educ_level"       
 [88] "partner_occupation"        "mother_occupation"         "mother_work_location"     
 [91] "survey_year"               "region_dhs.x"              "admin_region_name.x"      
 [94] "province_dhs.x"            "province_name.x"           "municipality_dhs.x"       
 [97] "municipality_name.x"       "barangay_name.x"           "latitude.x"               
[100] "longitude.x"               "altitude_gps.x"            "altitude_dem.x"           
[103] "cc_fips.x"                 "source.x"                  "urban_rural.x"            
[106] "datum.x"                   "region_dhs.y"              "admin_region_name.y"      
[109] "province_dhs.y"            "province_name.y"           "municipality_dhs.y"       
[112] "municipality_name.y"       "barangay_name.y"           "latitude.y"               
[115] "longitude.y"               "altitude_gps.y"            "altitude_dem.y"           
[118] "cc_fips.y"                 "source.y"                  "urban_rural.y"            
[121] "datum.y"                   "region_dhs"                "admin_region_name"        
[124] "province_dhs"              "province_name"             "municipality_dhs"         
[127] "municipality_name"         "barangay_name"             "latitude"                 
[130] "longitude"                 "altitude_gps"              "altitude_dem"             
[133] "cc_fips"                   "source"                    "urban_rural"              
[136] "datum"                     "number_of_siblings"        "manila"                   
[139] "post_policy"               "manila_post"              
  [1] "case_id"                   "cluster_id"                "household_id"             
  [4] "sample_weight"             "mother_age_current"        "residence_type"           
  [7] "household_size"            "household_u5_count"        "region"                   
 [10] "province"                  "municipality"              "wealth_index"             
 [13] "wealth_score"              "birth_index"               "birth_order"              
 [16] "is_twin"                   "birth_month"               "birth_year"               
 [19] "birth_date"                "child_sex"                 "child_alive"              
 [22] "age_at_death"              "child_age"                 "preceding_birth_interval" 
 [25] "succeeding_birth_interval" "delivery_place"            "birth_size"               
 [28] "birth_weight"              "received_bcg"              "bcg_day"                  
 [31] "bcg_month"                 "bcg_year"                  "received_dpt1"            
 [34] "dpt1_day"                  "dpt1_month"                "dpt1_year"                
 [37] "received_polio1"           "polio1_day"                "polio1_month"             
 [40] "polio1_year"               "received_dpt2"             "dpt2_day"                 
 [43] "dpt2_month"                "dpt2_year"                 "received_polio2"          
 [46] "polio2_day"                "polio2_month"              "polio2_year"              
 [49] "received_dpt3"             "dpt3_day"                  "dpt3_month"               
 [52] "dpt3_year"                 "received_polio3"           "polio3_day"               
 [55] "polio3_month"              "polio3_year"               "received_measles"         
 [58] "measles_day"               "measles_month"             "measles_year"             
 [61] "ever_vaccinated"           "diarrhea_recent"           "fever_recent"             
 [64] "cough_recent"              "total_children_ever"       "sons_at_home"             
 [67] "daughters_at_home"         "sons_died"                 "daughters_died"           
 [70] "births_last5"              "births_last1"              "births_this_month"        
 [73] "mother_age_first_birth"    "currently_pregnant"        "living_children"          
 [76] "living_plus_preg"          "living_plus_preg6"         "contraceptive_method"     
 [79] "contraceptive_intent"      "wanted_last_child"         "mother_educ_level"        
 [82] "mother_literate"           "mother_educ_attain"        "marital_status"           
 [85] "mother_age_first_sex"      "ideal_children"            "partner_educ_level"       
 [88] "partner_occupation"        "mother_occupation"         "mother_work_location"     
 [91] "survey_year"               "region_dhs.x"              "admin_region_name.x"      
 [94] "province_dhs.x"            "province_name.x"           "municipality_dhs.x"       
 [97] "municipality_name.x"       "barangay_name.x"           "latitude.x"               
[100] "longitude.x"               "altitude_gps.x"            "altitude_dem.x"           
[103] "cc_fips.x"                 "source.x"                  "urban_rural.x"            
[106] "datum.x"                   "region_dhs.y"              "admin_region_name.y"      
[109] "province_dhs.y"            "province_name.y"           "municipality_dhs.y"       
[112] "municipality_name.y"       "barangay_name.y"           "latitude.y"               
[115] "longitude.y"               "altitude_gps.y"            "altitude_dem.y"           
[118] "cc_fips.y"                 "source.y"                  "urban_rural.y"            
[121] "datum.y"                   "region_dhs"                "admin_region_name"        
[124] "province_dhs"              "province_name"             "municipality_dhs"         
[127] "municipality_name"         "barangay_name"             "latitude"                 
[130] "longitude"                 "altitude_gps"              "altitude_dem"             
[133] "cc_fips"                   "source"                    "urban_rural"              
[136] "datum"                     "number_of_siblings"        "manila"                   
[139] "post_policy"               "manila_post"              
  [1] "case_id"                   "cluster_id"                "household_id"             
  [4] "sample_weight"             "mother_age_current"        "residence_type"           
  [7] "household_size"            "household_u5_count"        "region"                   
 [10] "province"                  "municipality"              "wealth_index"             
 [13] "wealth_score"              "birth_index"               "birth_order"              
 [16] "is_twin"                   "birth_month"               "birth_year"               
 [19] "birth_date"                "child_sex"                 "child_alive"              
 [22] "age_at_death"              "child_age"                 "preceding_birth_interval" 
 [25] "succeeding_birth_interval" "delivery_place"            "birth_size"               
 [28] "birth_weight"              "received_bcg"              "bcg_day"                  
 [31] "bcg_month"                 "bcg_year"                  "received_dpt1"            
 [34] "dpt1_day"                  "dpt1_month"                "dpt1_year"                
 [37] "received_polio1"           "polio1_day"                "polio1_month"             
 [40] "polio1_year"               "received_dpt2"             "dpt2_day"                 
 [43] "dpt2_month"                "dpt2_year"                 "received_polio2"          
 [46] "polio2_day"                "polio2_month"              "polio2_year"              
 [49] "received_dpt3"             "dpt3_day"                  "dpt3_month"               
 [52] "dpt3_year"                 "received_polio3"           "polio3_day"               
 [55] "polio3_month"              "polio3_year"               "received_measles"         
 [58] "measles_day"               "measles_month"             "measles_year"             
 [61] "ever_vaccinated"           "diarrhea_recent"           "fever_recent"             
 [64] "cough_recent"              "total_children_ever"       "sons_at_home"             
 [67] "daughters_at_home"         "sons_died"                 "daughters_died"           
 [70] "births_last5"              "births_last1"              "births_this_month"        
 [73] "mother_age_first_birth"    "currently_pregnant"        "living_children"          
 [76] "living_plus_preg"          "living_plus_preg6"         "contraceptive_method"     
 [79] "contraceptive_intent"      "wanted_last_child"         "mother_educ_level"        
 [82] "mother_literate"           "mother_educ_attain"        "marital_status"           
 [85] "mother_age_first_sex"      "ideal_children"            "partner_educ_level"       
 [88] "partner_occupation"        "mother_occupation"         "mother_work_location"     
 [91] "survey_year"               "region_dhs.x"              "admin_region_name.x"      
 [94] "province_dhs.x"            "province_name.x"           "municipality_dhs.x"       
 [97] "municipality_name.x"       "barangay_name.x"           "latitude.x"               
[100] "longitude.x"               "altitude_gps.x"            "altitude_dem.x"           
[103] "cc_fips.x"                 "source.x"                  "urban_rural.x"            
[106] "datum.x"                   "region_dhs.y"              "admin_region_name.y"      
[109] "province_dhs.y"            "province_name.y"           "municipality_dhs.y"       
[112] "municipality_name.y"       "barangay_name.y"           "latitude.y"               
[115] "longitude.y"               "altitude_gps.y"            "altitude_dem.y"           
[118] "cc_fips.y"                 "source.y"                  "urban_rural.y"            
[121] "datum.y"                   "region_dhs"                "admin_region_name"        
[124] "province_dhs"              "province_name"             "municipality_dhs"         
[127] "municipality_name"         "barangay_name"             "latitude"                 
[130] "longitude"                 "altitude_gps"              "altitude_dem"             
[133] "cc_fips"                   "source"                    "urban_rural"              
[136] "datum"                     "number_of_siblings"        "manila"                   
[139] "post_policy"               "manila_post"              
  [1] "case_id"                   "cluster_id"                "household_id"             
  [4] "sample_weight"             "mother_age_current"        "residence_type"           
  [7] "household_size"            "household_u5_count"        "region"                   
 [10] "province"                  "municipality"              "wealth_index"             
 [13] "wealth_score"              "birth_index"               "birth_order"              
 [16] "is_twin"                   "birth_month"               "birth_year"               
 [19] "birth_date"                "child_sex"                 "child_alive"              
 [22] "age_at_death"              "child_age"                 "preceding_birth_interval" 
 [25] "succeeding_birth_interval" "delivery_place"            "birth_size"               
 [28] "birth_weight"              "received_bcg"              "bcg_day"                  
 [31] "bcg_month"                 "bcg_year"                  "received_dpt1"            
 [34] "dpt1_day"                  "dpt1_month"                "dpt1_year"                
 [37] "received_polio1"           "polio1_day"                "polio1_month"             
 [40] "polio1_year"               "received_dpt2"             "dpt2_day"                 
 [43] "dpt2_month"                "dpt2_year"                 "received_polio2"          
 [46] "polio2_day"                "polio2_month"              "polio2_year"              
 [49] "received_dpt3"             "dpt3_day"                  "dpt3_month"               
 [52] "dpt3_year"                 "received_polio3"           "polio3_day"               
 [55] "polio3_month"              "polio3_year"               "received_measles"         
 [58] "measles_day"               "measles_month"             "measles_year"             
 [61] "ever_vaccinated"           "diarrhea_recent"           "fever_recent"             
 [64] "cough_recent"              "total_children_ever"       "sons_at_home"             
 [67] "daughters_at_home"         "sons_died"                 "daughters_died"           
 [70] "births_last5"              "births_last1"              "births_this_month"        
 [73] "mother_age_first_birth"    "currently_pregnant"        "living_children"          
 [76] "living_plus_preg"          "living_plus_preg6"         "contraceptive_method"     
 [79] "contraceptive_intent"      "wanted_last_child"         "mother_educ_level"        
 [82] "mother_literate"           "mother_educ_attain"        "marital_status"           
 [85] "mother_age_first_sex"      "ideal_children"            "partner_educ_level"       
 [88] "partner_occupation"        "mother_occupation"         "mother_work_location"     
 [91] "survey_year"               "region_dhs.x"              "admin_region_name.x"      
 [94] "province_dhs.x"            "province_name.x"           "municipality_dhs.x"       
 [97] "municipality_name.x"       "barangay_name.x"           "latitude.x"               
[100] "longitude.x"               "altitude_gps.x"            "altitude_dem.x"           
[103] "cc_fips.x"                 "source.x"                  "urban_rural.x"            
[106] "datum.x"                   "region_dhs.y"              "admin_region_name.y"      
[109] "province_dhs.y"            "province_name.y"           "municipality_dhs.y"       
[112] "municipality_name.y"       "barangay_name.y"           "latitude.y"               
[115] "longitude.y"               "altitude_gps.y"            "altitude_dem.y"           
[118] "cc_fips.y"                 "source.y"                  "urban_rural.y"            
[121] "datum.y"                   "region_dhs"                "admin_region_name"        
[124] "province_dhs"              "province_name"             "municipality_dhs"         
[127] "municipality_name"         "barangay_name"             "latitude"                 
[130] "longitude"                 "altitude_gps"              "altitude_dem"             
[133] "cc_fips"                   "source"                    "urban_rural"              
[136] "datum"                     "number_of_siblings"        "manila"                   
[139] "post_policy"               "manila_post"