### 2025-08-07 data-analysis/outputs/summary_table_outcomes.tex

\begingroup\fontsize{7.5}{9.5}\selectfont

\begin{longtable}[t]{>{}l>{\centering\arraybackslash}p{1.6cm}>{\centering\arraybackslash}p{1.6cm}>{\centering\arraybackslash}p{1.6cm}>{\centering\arraybackslash}p{1.6cm}>{\centering\arraybackslash}p{1.6cm}>{\centering\arraybackslash}p{1.6cm}}
\caption{Descriptive statistics by location and policy period (outcomes only).}\\
\toprule
\multicolumn{1}{c}{ } & \multicolumn{6}{c}{Group} \\
\cmidrule(l{3pt}r{3pt}){2-7}
Variable & Manila, Pre-Policy  N = 209 & Manila, Post-Policy  N = 165 & Other NCR, Pre-Policy  N = 123 & Other NCR, Post-Policy  N = 136 & Non-NCR, Pre-Policy  N = 7,437 & Non-NCR, Post-Policy  N = 3,410\\
\midrule
\endfirsthead
\caption[]{Descriptive statistics by location and policy period (outcomes only). \textit{(continued)}}\\
\toprule
\multicolumn{1}{c}{ } & \multicolumn{6}{c}{Group} \\
\cmidrule(l{3pt}r{3pt}){2-7}
Variable & Manila, Pre-Policy  N = 209 & Manila, Post-Policy  N = 165 & Other NCR, Pre-Policy  N = 123 & Other NCR, Post-Policy  N = 136 & Non-NCR, Pre-Policy  N = 7,437 & Non-NCR, Post-Policy  N = 3,410\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\textbf{Birth weight (grams)} & 4,766.51 (3,089.10) & 3,864.21 (2,419.37) & 5,816.11 (3,592.43) & 5,379.28 (3,411.51) & 5,754.47 (3,399.87) & 5,254.08 (3,321.64)\\
\textbf{Perceived birth size} &  &  &  &  &  & \\
\hspace{1em}1 - Very small & 10 (4.8\%) & 1 (0.6\%) & 2 (1.6\%) & 0 (0\%) & 345 (4.7\%) & 116 (3.4\%)\\
\hspace{1em}2 - Smaller than average & 35 (17\%) & 39 (24\%) & 25 (20\%) & 34 (25\%) & 1,256 (17\%) & 584 (17\%)\\
\hspace{1em}3 - Average size & 122 (59\%) & 92 (56\%) & 76 (62\%) & 84 (62\%) & 4,397 (59\%) & 1,990 (59\%)\\
\addlinespace
\hspace{1em}4 - Larger than average & 28 (13\%) & 22 (13\%) & 18 (15\%) & 15 (11\%) & 970 (13\%) & 509 (15\%)\\
\hspace{1em}5 - Very large & 12 (5.8\%) & 11 (6.7\%) & 2 (1.6\%) & 3 (2.2\%) & 407 (5.5\%) & 184 (5.4\%)\\
\hspace{1em}8 - Don't know & 0 (0\%) & 0 (0\%) & 0 (0\%) & 0 (0\%) & 13 (0.2\%) & 14 (0.4\%)\\
\hspace{1em}9 - Missing & 1 (0.5\%) & 0 (0\%) & 0 (0\%) & 0 (0\%) & 18 (0.2\%) & 0 (0\%)\\
\textbf{Total number of vaccines received} &  &  &  &  &  & \\
\addlinespace
\hspace{1em}1–2 Vaccines & 1 (0.5\%) & 3 (1.8\%) & 1 (0.8\%) & 3 (2.2\%) & 68 (0.9\%) & 95 (2.8\%)\\
\hspace{1em}3–4 Vaccines & 4 (1.9\%) & 5 (3.0\%) & 3 (2.4\%) & 5 (3.7\%) & 136 (1.8\%) & 117 (3.4\%)\\
\hspace{1em}5+ Vaccines & 38 (18\%) & 33 (20\%) & 31 (25\%) & 52 (38\%) & 1,869 (25\%) & 1,141 (33\%)\\
\hspace{1em}No Vaccines & 166 (79\%) & 124 (75\%) & 88 (72\%) & 76 (56\%) & 5,364 (72\%) & 2,057 (60\%)\\*
\multicolumn{7}{l}{\rule{0pt}{1em}\textsuperscript{1} Mean (SD); n (\%)}\\
\end{longtable}
\endgroup{}