### 2025-08-07 data-analysis/scripts/98c_paper_outputs.qmd
add scripts for generating paper output tables
Introduces three new scripts: 98a_paper_outputs.R for summary tables of birth weight, perceived birth size, and vaccine categories; 98b_paper_outputs.R for census-based birth rate calculations and LaTeX export; and 98c_paper_outputs.qmd for first-stage regression analysis and LaTeX table output. These scripts automate the creation of key tables for the paper using cleaned data sources.

---
title: "98c_paper_outputs"
format: pdf
---

```{r}
# Core data wrangling
library(dplyr)
library(stats)
library(MASS)
library(broom)
library(modelsummary)   
library(ggplot2)

```

```{r}
dhs1 <- readRDS("../clean-data/dhs_kr_children_with_geo.rds") |>
  filter(survey_year %in% c(1998, 2003, 2008)) |>
    mutate(
      number_of_siblings = living_children - 1)

dhs2 <- dhs1 %>%
  mutate(
    manila = as.character(municipality_name) == "Manila",
    post_policy = birth_year >= 2000,
    manila_post = manila * post_policy
  )
```

```{r}
library(fixest)
library(kableExtra)

# 1. First-stage: total children ever
first_stage1 <- feols(
  total_children_ever ~ manila + post_policy + manila_post +
    mother_educ_attain + mother_age_current + wealth_index | survey_year,
  data = dhs2
)

# 2. First-stage: number of living children
first_stage2 <- feols(
  living_children ~ manila + post_policy + manila_post +
    mother_educ_attain + mother_age_current + wealth_index | survey_year,
  data = dhs2
)

# Name your models
models <- list(
  "Total Children Ever" = first_stage1,
  "Living Children" = first_stage2
)

# Prepare LaTeX table string
table_latex <- modelsummary(
  models,
  output = "latex",
  statistic = "p.value",
  coef_map = c(
    manila = "Manila",
    post_policy = "Post-Policy",
    manila_post = "Manila × Post",
    mother_educ_attain = "Mother's Education",
    mother_age_current = "Mother's Age",
    wealth_index = "Wealth Index"
  ),
  gof_omit = "IC|Log|Std.Errors",
  notes = list("All regressions include survey year fixed effects."),
  fmt = 3,
  stars = TRUE,
  escape = FALSE,
  latex_options = "tabular"  # prevents talltblr error
)

# Save to .tex file
writeLines(
  capture.output(print(table_latex)),
  con = "../outputs/first_stage_sibship.tex"
)

```

```{r}
names(dhs2)
```